---
title: "ACTIONet Pan-Cancer"
output: html_notebook
---

```{r}
library(Seurat)
library(ACTIONet)
library(SingleCellExperiment)
library(tidyverse)
library(viridis)
library(RColorBrewer)
library(UpSetR)
library(ComplexHeatmap)
```

# Load the data
```{r}
seurat <- readRDS("../data/pan_cancer_epi.rds")
```

```{r}
DimPlot(seurat, group.by="Cancer")
```

```{r}
table(seurat$SampleID)
```


```{r}
#Number of unique samples
length(unique(seurat$SampleID))
```

## Get sample list
```{r}
sample_list <- unique(seurat$SampleID)
```


# ACTIONet
Subset seurat object by sample and run ACTIONet on it
```{r}
runACTION <- function(sample){
  print(paste0("Processing sample: ", sample))
  #Subset seurat and convert to SCE
  cells_keep <- colnames(seurat)[seurat$SampleID == sample]
  seurat_subset <- CreateSeuratObject(counts = seurat[["RNA"]]@counts[,cells_keep],
                                      meta.data = seurat@meta.data[cells_keep,],
                                      min.cells=3)
  seurat_subset <- NormalizeData(seurat_subset)
  sce <- Seurat::as.SingleCellExperiment(seurat_subset, assay="RNA")
  
  #Run ACTIONet
  ace <- reduce.ace(sce, reduced_dim = 25) #As far as I can tell, this relates to dimensionality of the data. 50 is default. Works well for full tissue, but I think it leads to representing some noise on these pure populations
  rm(sce)
  rm(seurat_subset)
  action <- run.ACTIONet(ace=ace, k_max=6,
                         min_cells_per_arch = 5) #Tried some tweaks and this works pretty well for me
  return(action)
}
```

```{r}
action_list <- lapply(sample_list, runACTION)
names(action_list) <- sample_list

saveRDS(action_list, file="../output/ACTIONet_list.rds")
```

```{r}
action_list <- readRDS("../output/ACTIONet_list.rds")
```


# Build master metadata
```{r}
meta_list <- list()
for(i in 1:length(action_list)){
  meta_list[[i]] <- data.frame(Dim1 = action_list[[i]]$ACTIONet2D[,1],
                               Dim2 = action_list[[i]]$ACTIONet2D[,2],
                               Archetype_Membership = factor(action_list[[i]]$assigned_archetype))
  colnames(action_list[[i]]$archetype_footprint) <- paste0("A", 1:ncol(action_list[[i]]$archetype_footprint))
  meta_list[[i]] <- cbind(meta_list[[i]], action_list[[i]]$archetype_footprint)
}
names(meta_list) <- sample_list
```


# EMT Gene Set Scoring
## Load gene sets
```{r}
hallmarks <- fgsea::gmtPathways("~/Data/GeneLists/hallmark.genesets.v6.1.symbols.gmt")
hallmark_emt <- hallmarks["HALLMARK_EPITHELIAL_MESENCHYMAL_TRANSITION"]

go_terms <- fgsea::gmtPathways("~/Data/GeneLists/GOTerms.BP.v6.1.symbols.gmt")
go_emt <- go_terms["GO_EPITHELIAL_TO_MESENCHYMAL_TRANSITION"]

#cancerSEA
cancer_sea <- readRDS("~/Data/GeneLists/cancerSEA/cancerSEA_gene_sets.rds")
names(cancer_sea) <- paste0("CancerSEA_", names(cancer_sea))
cancer_sea_emt <- cancer_sea["CancerSEA_EMT"]

#My conserved upregulated list
conserved_mes <- list(readr::read_lines("~/Projects/emt_dynamics/output/conserved_upregulated_genes.csv")[-1])
names(conserved_mes) <- "Conserved_EMT_Up"

#My conserved downregulated list
conserved_epi <- list(readr::read_lines("~/Projects/emt_dynamics/output/conserved_downregulated_genes.csv")[-1])
names(conserved_epi) <- "Conserved_EMT_Down"

#dbEMT
dbEMT <- read.csv("~/Data/GeneLists/Zhao_dbEMT.csv")
dbEMT <- list(dbEMT$Symbol)
names(dbEMT) <- "dbEMT"

#Puram HNSCC pEMT
puram <- read.csv("~/Data/GeneLists/puram_hnscc_pEMT_signatures.csv")
puram <- list(unique(puram$Gene)) #All unique genes from their list of 50-gene NMF programs
names(puram) <- "Puram_pEMT"

#Taube Core EMT Up
taube_up <- list(read_lines("~/Data/GeneLists/Taube_Core_EMT_Up.txt"))
names(taube_up) <- "Taube_Core_EMT_Up"

#Kinker et al. EMT signatures
kinker <- read.csv("~/Data/GeneLists/Kinker_Tirosh_RHPs.csv")
kinker_1 <- list(Kinker_1 = kinker$EMT.I)
kinker_2 <- list(Kinker_2 = kinker$EMT.II)
kinker_3 <- list(Kinker_3 = kinker$EMT.III)
```

```{r}
emt_sets <- c(cancer_sea_emt, hallmark_emt,
             go_emt,  conserved_mes, dbEMT, puram, taube_up,
             kinker_1, kinker_2, kinker_3)
```

```{r}
names(emt_sets)
```

```{r}
names(emt_sets) <- c("CancerSEA", "Hallmark", "GO", 
                     "Cook", "dbEMT", "Puram_pEMT", "Taube",
                     "Kinker_I", "Kinker_II", "Kinker_III")
```

## Compare composition
### UpSet Plot
```{r}
library(UpSetR)
```

```{r}
emt_genes <- unique(unlist(emt_sets))
emt_membership <- fromList(emt_sets)
rownames(emt_membership) <- emt_genes
```

```{r}
upset_plot <- upset(emt_membership, order.by="freq",
      nsets = length(emt_sets),
      nintersects = 15,
      text.scale = c(1.3, 1.3, 1.3, 1.3, 1.5, 1.3),
      mb.ratio = c(0.4, 0.6))
```

```{r}
cairo_pdf(filename = "../figs/emt_geneset_upset.pdf",
          width=5, height=4)
upset_plot
dev.off()
```

### Set count
```{r}
list_counts <- as.data.frame(rowSums(emt_membership))
colnames(list_counts) <- "Count"
list_counts$Gene <- rownames(emt_membership)
list_counts <- arrange(list_counts, desc(Count))
list_counts$Index <- 1:nrow(list_counts)
```

```{r}
count_plot <- ggplot(list_counts, aes(x=Index, y=Count)) +
  geom_point(size=0.25, color="black") +
  geom_vline(xintercept = 0, linetype = 2, size=0.25) +
  xlab("") + ylab("Gene set count") +
  scale_y_continuous(limits = c(1,10), breaks = c(2,4,6,8,10)) +
  theme_classic() +
  theme(axis.text=element_text(size=5, color="black"),
        axis.title=element_text(size=5),
        axis.line = element_line(size=0.25),
        axis.ticks = element_line(size=0.25))
ggsave(count_plot, filename = "../figs/EMT_gene_set_frequency.pdf",
       width=3.2, height=3, units="cm")
```

```{r}
count_plot
```

```{r}
list_counts
```

## Scoring
We'll just run Seurat's AddModuleScore on each individual sample
```{r}
for(i in 1:length(sample_list)){
  seurat_subset <- subset(seurat, subset = SampleID == sample_list[i])
  seurat_subset <- CellCycleScoring(seurat_subset, s.features = cc.genes.updated.2019$s.genes,
                                    g2m.features = cc.genes.updated.2019$g2m.genes)
  # Gene set score each EMT gene set
  for(n in 1:length(emt_sets)){
    seurat_subset <- AddModuleScore(seurat_subset, features = emt_sets[n],
                                    name = names(emt_sets)[n])
  }
  
  colnames(seurat_subset@meta.data)[87:96] <- gsub("1", "", colnames(seurat_subset@meta.data)[87:96]) #remove the stupid 1
  
  #Merge data into meta_list
  meta_list[[i]] <- cbind(meta_list[[i]], 
                        seurat_subset@meta.data[,c(names(emt_sets),  
                                                   "S.Score", "G2M.Score", "Phase", 
                                                   "percent.mito", "nCount_RNA")])
}

saveRDS(meta_list, file = "../output/master_metadata.rds")
```


# Median EMT Score
```{r}
master_meta <- do.call("bind_rows", meta_list)
master_meta$SampleID <- seurat$SampleID[rownames(master_meta)]
```

```{r}
emt_avg <- master_meta[,c("SampleID", names(emt_sets))]
emt_avg <- tidyr::pivot_longer(emt_avg,
                               !SampleID,
                               names_to="gene_set",
                               values_to="score")
emt_avg <- emt_avg %>% 
  group_by(SampleID, gene_set) %>%
  summarise(median = median(score))

emt_avg <- emt_avg %>% pivot_wider(names_from=gene_set, values_from=median)
```

```{r}
emt_mat <- as.matrix(emt_avg[,2:ncol(emt_avg)])
rownames(emt_mat) <- emt_avg$SampleID
emt_mat <- scale(emt_mat, scale=T, center=T)
#emt_mat[emt_mat > 1.5] <- 1.5
#emt_mat[emt_mat < -1.5] <- -1.5
```



```{r}
cols <- paletteer::paletteer_c("scico::berlin", n=100)

pdf(file = "../figs/emt_geneset_score_median.pdf",
    width=4, height=12.5)
pheatmap(emt_mat,
         color = cols,
         breaks = seq(-2, 2, length.out = 100),
         border_color="black",
         show_colnames=T,
         angle_col = "45",
         fontsize_row = 4.5,
         treeheight_col = 15,
         treeheight_row = 15,
         legend=F,
         name = NULL,
         clustering_method="ward.D2")
dev.off()

#Just to get berlin legend:
pheatmap::pheatmap(emt_mat,
         color = cols,
         breaks = seq(-2, 2, length.out = 101),
         border_color="black",
         show_colnames=T,
         angle_col = "45",
         treeheight_col = 10,
         treeheight_row = 10,
         fontsize_row = 4.5,
         legend=T,
         name = NULL,
         clustering_method="ward.D2",
         file = "../figs/emt_geneset_score_median_4Legend.png",
    width=4, height=12.5)
```

# Correlation of EMT gene sets to each other
```{r}
emt_cor <- cor(master_meta[,names(emt_sets)])
```

```{r}
pdf(file = "../figs/emt_geneset_score_correlation.pdf",
    width=4.25, height=3.2)
pheatmap(emt_cor,
         color = viridis::inferno(100),
         #breaks = seq(0, 1, length.out=101),
         show_colnames=F,
         treeheight_col = 0,
         treeheight_row = 15,
         legend=F,
         fontsize_row = 12,
         clustering_method="ward.D2",
         border_color = "black",
         display_numbers = T,
         number_color="grey50")
dev.off()
```

# Plot archetype membership
## By ACTIONet embedding
```{r}
cols <- RColorBrewer::brewer.pal(n=9, "Set1")
cols <- cols[-6]
```

```{r}
for(i in 1:length(sample_list)){
  arch_plot <- ggplot(meta_list[[i]], aes(x=Dim1, y=Dim2)) +
    geom_point(size=1, shape=16, aes(color=Archetype_Membership)) +
    scale_color_manual(values = cols[1:length(unique(meta_list[[i]]$Archetype_Membership))]) +
    theme_classic() +
    theme(axis.text = element_text(size=12, color="black"),
          legend.text = element_text(size=12))
  ggsave(arch_plot, device = cairo_pdf(),
         filename = paste0("../figs/archetype_membership/",
                           sample_list[[i]], "_arch_membership.pdf"),
         width=5.75, height=3.5)
  dev.off()
}
```

## By PCA
```{r}
plotArch_PCA <- function(sampleID){
  print(paste0("Processing sample: ", sampleID))
  cells <- rownames(meta_list[[sampleID]])
  seurat_subset <- subset(seurat, cells = cells)
  seurat_subset <- SCTransform(seurat_subset, vars.to.regress="percent.mito")
  seurat_subset <- RunPCA(seurat_subset, verbose=F)
  seurat_subset <- RunUMAP(seurat_subset, dims=1:30)
  
  dat <- meta_list[[sampleID]]
  dat$PC1 <- Embeddings(seurat_subset, "pca")[,1]
  dat$PC2 <- Embeddings(seurat_subset, "pca")[,2]
  
  #Get specific cell for each archetype
  dat$Archetype_Membership <- paste0("A", dat$Archetype_Membership)
  archs <- unique(dat$Archetype_Membership)
  arch_dat <- dat[,c("Archetype_Membership", archs)]
  arch_dat$Cell <- rownames(dat)
  arch_dat <- arch_dat %>% 
    pivot_longer(!c(Archetype_Membership, Cell), 
                 names_to = "Archetype",
                 values_to = "Score") %>%
    group_by(Archetype) %>%
    filter(Score == max(Score))
  arch_dat$PC1 <- dat[arch_dat$Cell, "PC1"]
  arch_dat$PC2 <- dat[arch_dat$Cell, "PC2"]
  
  
  pca_plot <- ggplot(dat, aes(PC1, PC2)) +
    geom_point(size=1, shape=16, aes(color=Archetype_Membership), alpha=0.5) +
    geom_point(data=arch_dat, aes(fill=Archetype_Membership), shape=21, 
               colour="black", size=4) +
    scale_color_brewer(palette="Dark2") +
    scale_fill_brewer(palette="Dark2") +
    theme_classic() +
    theme(axis.text=element_text(size=10, color="black"),
          axis.title=element_text(size=12),
          legend.title=element_text(size=12),
          legend.text=element_text(size=12))
  
  ggsave(pca_plot, filename= paste0("../figs/archetype_membership_pca/",
                                    sampleID,  "_arch_membership.pdf"),
         width=4.5, height=2.5)
  
}


lapply(sample_list, plotArch_PCA)
```


# Correlation of Archetypes w/ EMT

```{r}
getCor <- function(df){
  total_arch <- unique(paste0("A", unique(df$Archetype_Membership)))
  total_arch <- total_arch[order(total_arch)]
  cor_mat <- cor(df[,total_arch], df[,names(emt_sets)])
}

cor_list <- lapply(meta_list, getCor)
```

Need to get a list that matches source to sample

```{r}
annotation <- unique(seurat@meta.data[,c("Source", "Cancer", "SampleID")])
annotation$Source <- paste0(annotation$Source, "_", annotation$Cancer)
unique_source <- unique(annotation$Source)
```

Go source by source, making heatmap lists for each patient within that source

Oops, just meant to have gene sets ordered the same in each list

```{r}
gene_set_order <- c("Cook", "Puram_pEMT", "Kinker_III", "dbEMT", "Kinker_I",
                    "GO", "CancerSEA", "Taube", "Hallmark", "Kinker_II")
```


```{r}
all_heatmaps <- list()
for(i in 1:length(unique_source)){
  samples <- annotation %>% filter(Source == unique_source[i]) %>% pull(SampleID)
  
  heat_list <- NULL
  for(n in 1:length(samples)){
    heat <- ComplexHeatmap::pheatmap(t(cor_list[[samples[n]]])[gene_set_order,],
                                   breaks=seq(-0.5,0.5, length.out=100),
                                   color = colorRampPalette(rev(brewer.pal(n = 7, name = "RdBu")))(100),
                                  cluster_rows=F,
                                  cluster_cols=T,
                                  clustering_method="ward.D2",
                                  treeheight_col = 0,
                                  fontsize = 10,
                                  legend = F,
                                  border_col="black")
    heat_list <- heat_list + heat
  }
  all_heatmaps[[i]] <- heat_list
  draw(all_heatmaps[[i]])
}
```

```{r}
table(annotation$Source)
```

```{r}
unique_source #order of all_heatmaps
```


```{r}
pdf(file = "../figs/archetype_emt_correlation_Durante_Uveal.pdf",
    width=14, height=2.15)
all_heatmaps[[1]]
dev.off()

pdf(file = "../figs/archetype_emt_correlation_Kim_Lung.pdf",
    width=19, height=2.15)
all_heatmaps[[2]]
dev.off()

pdf(file = "../figs/archetype_emt_correlation_Lambrechts_Lung.pdf",
    width=9.5, height=2.15)
all_heatmaps[[3]]
dev.off()

pdf(file = "../figs/archetype_emt_correlation_Laughney_Lung.pdf",
    width=8.35, height=2.15)
all_heatmaps[[4]]
dev.off()

pdf(file = "../figs/archetype_emt_correlation_Lee_Colorectal.pdf",
    width=23.5, height=2.15)
all_heatmaps[[5]]
dev.off()

pdf(file = "../figs/archetype_emt_correlation_Ji_SCC.pdf",
    width=6.2, height=2.15)
all_heatmaps[[6]]
dev.off()

pdf(file = "../figs/archetype_emt_correlation_Starr_Ovarian.pdf",
    width=5.5, height=2.15)
all_heatmaps[[7]]
dev.off()

pdf(file = "../figs/archetype_emt_correlation_Steele_PDAC.pdf",
    width=14.5, height=2.15)
all_heatmaps[[8]]
dev.off()

pdf(file = "../figs/archetype_emt_correlation_Uhlitz_Colorectal.pdf",
    width=6.25, height=2.15)
all_heatmaps[[9]]
dev.off()

pdf(file = "../figs/archetype_emt_correlation_Wu_Breast.pdf",
    width=4.8, height=2.15)
all_heatmaps[[10]]
dev.off()

pdf(file = "../figs/archetype_emt_correlation_Sathe_Gastric.pdf",
    width=8, height=2.15)
all_heatmaps[[11]]
dev.off()

pdf(file = "../figs/archetype_emt_correlation_Qian_Breast.pdf",
    width=16, height=2.15)
all_heatmaps[[12]]
dev.off()

pdf(file = "../figs/archetype_emt_correlation_Qian_Colorectal.pdf",
    width=15, height=2.15)
all_heatmaps[[13]]
dev.off()

pdf(file = "../figs/archetype_emt_correlation_Qian_Lung.pdf",
    width=19, height=2.15)
all_heatmaps[[14]]
dev.off()

pdf(file = "../figs/archetype_emt_correlation_Qian_Ovarian.pdf",
    width=7, height=2.15)
all_heatmaps[[15]]
dev.off()
```


# EMT correlation values
```{r}
cor_list_merge <- cor_list
for(i in 1:length(cor_list_merge)){
  rownames(cor_list_merge[[i]]) <- paste0(names(cor_list_merge)[i], "_", rownames(cor_list_merge[[i]]))
}
cor_list_merge <- do.call("rbind", cor_list_merge)
```

```{r}
write.csv(cor_list_merge, file="../output/archetype_emt_correlation.csv")
```



