---
title: "Combine epithelial cells"
output: html_notebook
---

# Dependencies

```{r}
library(Seurat)
library(Matrix)
library(tidyverse)
library(ggrastr)
```
# General functions
Sometimes you get stray cells annotated as "epithelial", but reside in a cluster that is very clearly some other cell type (eg. T cell). This may cause some weird results when we subset in on cells from individual patients. **These are the vast minority**. To clean this issue up, we'll only consider cells 
1) Annotated as "Epithelial_cells" from singleR
2i) Reside in a cluster that contains >1% of all epithelial cells (eg. if 5/2000 of all epithelial cells in a specific cluster, we remove them)
2ii) Reside in a cluster where "Epithelial" is the most abundant annotation (in case there's a tiny cluster that isn't >1%, but is dominantly epithelial)

Filter on (1) and (2i OR 2ii)

```{r}
# 2i) Which clusters have >1% of all the epithelial cells
getHighFreqClusters <- function(seurat){
  epi_counts_threshold <- seurat@meta.data %>% 
    filter(CellType == "Epithelial_cells") %>% 
    group_by(seurat_clusters) %>%
    summarise(count = n()) %>%
    mutate(prop = count / sum(count)) %>%
    filter(prop >= 0.01) %>%
    pull(seurat_clusters) %>%
    as.character()
}
```

```{r}
# 2iI) Which clusters are dominantly epithelial
getDominantClusters <- function(seurat){
  epi_clusters <- seurat@meta.data %>% group_by(seurat_clusters, CellType) %>%
    summarise(count = n()) %>%
    mutate(prop = count / sum(count)) %>%
    top_n(1, prop) %>%
    filter(CellType == "Epithelial_cells") %>%
    pull(seurat_clusters) %>%
    as.character()
}

```

```{r}
getEpithelial <- function(seurat){
  high_freq_clusters <- getHighFreqClusters(seurat)
  epi_dominant_clusters <- getDominantClusters(seurat)
  
  epi_meta <- seurat@meta.data %>%
    filter(CellType == "Epithelial_cells") %>%
    filter(seurat_clusters %in% high_freq_clusters |
             seurat_clusters %in% epi_dominant_clusters)
  
  ###Summary of filter
  raw_epi <- seurat@meta.data %>% filter(CellType == "Epithelial_cells") %>% nrow()
  filtered_epi <- nrow(epi_meta)
  print(paste0(filtered_epi, "/", raw_epi, " Epithelial cells kept"))
  
  DefaultAssay(seurat) <- "RNA"
  seurat <- subset(DietSeurat(seurat), cells = rownames(epi_meta))
  return(seurat)
}
```


# Colorectal - Lee et al - GSE144735 (SMC)

```{r}
seurat <- readRDS("../output/tumour_annotations/colorectal_lee_seurat.rds")
#Add some metadata to help post-merge
seurat$Cancer <-  "Colorectal"
seurat$Source <- "Lee"
```

```{r}
DimPlot(seurat, group.by="CellType", label=T)
DimPlot(seurat, group.by="seurat_clusters", label=T)
```

```{r}
colorectal_lee <- getEpithelial(seurat)
```

# Colorectal - Lee et al - GSE132465 (KUL)
```{r}
seurat <- readRDS("../output/tumour_annotations/colorectal_lee_KUL_seurat.rds")
#Add some metadata to help post-merge
seurat$Cancer <-  "Colorectal"
seurat$Source <- "Lee"
```

```{r}
DimPlot(seurat, group.by="CellType", label=T)
DimPlot(seurat, group.by="seurat_clusters", label=T)
```

```{r}
colorectal_lee_kul <- getEpithelial(seurat)
```

# Colorectal - Uhlitz et al - Data direct from authors (preprint)
```{r}
seurat <- readRDS("../output/tumour_annotations/colorectal_Uhlitz_seurat.rds")
#Add some metadata to help post-merge
seurat$Cancer <-  "Colorectal"
seurat$Source <- "Uhlitz"
```

```{r}
DimPlot(seurat, group.by="CellType", label=T)
DimPlot(seurat, group.by="seurat_clusters", label=T)
```

```{r}
colorectal_uhlitz <- getEpithelial(seurat)
```

```{r}
colorectal_uhlitz$Patient <- colorectal_uhlitz$orig.ident
```


# Colorectal - Qian et al - http://blueprint.lambrechtslab.org

```{r}
seurat <- readRDS("../output/tumour_annotations/colorectal_Qian_seurat.rds")
#Add some metadata to help post-merge
seurat$Cancer <-  "Colorectal"
seurat$Source <- "Qian"
```

```{r}
DimPlot(seurat, group.by="CellType", label=T)
DimPlot(seurat, group.by="seurat_clusters", label=T)
```

```{r}
colorectal_qian <- getEpithelial(seurat)
```

```{r}
colorectal_qian$Patient <- colorectal_qian$orig.ident
```

# Gastric - Sathe et al - https://dna-discovery.stanford.edu

```{r}
seurat <- readRDS("../output/tumour_annotations/gastric_sathe_seurat.rds")
#Add some metadata to help post-merge
seurat$Cancer <-  "Gastric"
seurat$Source <- "Sathe"
```

```{r}
DimPlot(seurat, group.by="CellType", label=T)
DimPlot(seurat, group.by="seurat_clusters", label=T)
```

```{r}
gastric_sathe <- getEpithelial(seurat)
```

```{r}
gastric_sathe$Patient <- gastric_sathe$orig.ident
```

# Lung - Kim et al - GSE131907

```{r}
seurat <- readRDS("../output/tumour_annotations/lung_kim_seurat.rds")
#Add some metadata to help post-merge
seurat$Cancer <-  "Lung"
seurat$Source <- "Kim"
seurat$Patient <- seurat$Sample
```

```{r}
DimPlot(seurat, group.by="CellType", label=T)
DimPlot(seurat, group.by="seurat_clusters", label=T)
```

```{r}
lung_kim <- getEpithelial(seurat)
```

```{r}
lung_kim$Patient <- lung_kim$Sample
```

# Lung - Lambrechts et al - E-MTAB-6149 & E-MTAB-6653

```{r}
seurat <- readRDS("../output/tumour_annotations/lung_lambrechts_seurat.rds")
#Add some metadata to help post-merge
seurat$Cancer <-  "Lung"
seurat$Source <- "Lambrechts"
```

```{r}
DimPlot(seurat, group.by="CellType", label=T)
DimPlot(seurat, group.by="seurat_clusters", label=T)
```

```{r}
lung_lambrechts <- getEpithelial(seurat)
```

```{r}
lung_lambrechts$Patient <- lung_lambrechts$orig.ident
```

# Lung - Laughey et al - GSE123904

```{r}
seurat <- readRDS("../output/tumour_annotations/lung_laughney_seurat.rds")
#Add some metadata to help post-merge
seurat$Cancer <-  "Lung"
seurat$Source <- "Laughney"
```

```{r}
DimPlot(seurat, group.by="CellType", label=T)
DimPlot(seurat, group.by="seurat_clusters", label=T)
```

```{r}
lung_laughney <- getEpithelial(seurat)
```


# Lung - Qian et al - http://blueprint.lambrechtslab.org

```{r}
seurat <- readRDS("../output/tumour_annotations/lung_qian_seurat.rds")
#Add some metadata to help post-merge
seurat$Cancer <-  "Lung"
seurat$Source <- "Qian"
```

```{r}
DimPlot(seurat, group.by="CellType", label=T)
DimPlot(seurat, group.by="seurat_clusters", label=T)
```

```{r}
lung_qian <- getEpithelial(seurat)
```

# Lung - Wu et al - GSE148071

```{r}
seurat <- readRDS("../output/tumour_annotations/lung_wu_seurat.rds")
#Add some metadata to help post-merge
seurat$Cancer <-  "Lung"
seurat$Source <- "Wu"
```

```{r}
DimPlot(seurat, group.by="CellType", label=T)
DimPlot(seurat, group.by="seurat_clusters", label=T)
```

```{r}
lung_wu <- getEpithelial(seurat)
```

```{r}
lung_wu$Patient <- lung_wu$orig.ident
```

# Ovarian - Geistlinger et al - GSE154600

```{r}
seurat <- readRDS("../output/tumour_annotations/ovarian_geistlinger_seurat.rds")
#Add some metadata to help post-merge
seurat$Cancer <-  "Ovarian"
seurat$Source <- "Geistlinger"
```

```{r}
DimPlot(seurat, group.by="CellType", label=T)
DimPlot(seurat, group.by="seurat_clusters", label=T)
```

```{r}
ovarian_geistlinger <- getEpithelial(seurat)
```

```{r}
ovarian_geistlinger$Patient <- ovarian_geistlinger$orig.ident
```

# Ovarian - Qian et al - http://blueprint.lambrechtslab.org

```{r}
seurat <- readRDS("../output/tumour_annotations/ovarian_qian_seurat.rds")
#Add some metadata to help post-merge
seurat$Cancer <-  "Ovarian"
seurat$Source <- "Qian"
```

```{r}
DimPlot(seurat, group.by="CellType", label=T)
DimPlot(seurat, group.by="seurat_clusters", label=T)
```

```{r}
ovarian_qian <- getEpithelial(seurat)
```


# PDAC - Steele et al - GSE155698

```{r}
seurat <- readRDS("../output/tumour_annotations/pdac_steele_seurat.rds")
#Add some metadata to help post-merge
seurat$Cancer <-  "PDAC"
seurat$Source <- "Steele"
```

```{r}
DimPlot(seurat, group.by="CellType", label=T)
DimPlot(seurat, group.by="seurat_clusters", label=T)
```

```{r}
pdac_steele <- getEpithelial(seurat)
```

```{r}
pdac_steele$Patient <- gsub("PDAC_", "", pdac_steele$Sample)
```

# Breast - Wu et al - ENA PRJEB35405

```{r}
seurat <- readRDS("../output/tumour_annotations/breast_wu_seurat.rds")
#Add some metadata to help post-merge
seurat$Cancer <-  "Breast"
seurat$Source <- "Wu"
```

```{r}
DimPlot(seurat, group.by="CellType", label=T)
DimPlot(seurat, group.by="seurat_clusters", label=T)
```

```{r}
breast_wu <- getEpithelial(seurat)
```

# Breast - Qian et al - http://blueprint.lambrechtslab.org

```{r}
seurat <- readRDS("../output/tumour_annotations/breast_qian_seurat.rds")
#Add some metadata to help post-merge
seurat$Cancer <-  "Breast"
seurat$Source <- "Qian"
```

```{r}
DimPlot(seurat, group.by="CellType", label=T)
DimPlot(seurat, group.by="seurat_clusters", label=T)
```

```{r}
breast_qian <- getEpithelial(seurat)
```



# Breast - Bassez et al - http://biokey.lambrechtslab.org/

## Cohort 1

```{r}
seurat <- readRDS("../output/tumour_annotations/breast_bassez_cohort1_seurat.rds")
#Add some metadata to help post-merge
seurat$Cancer <-  "Breast"
seurat$Source <- "Bassez"
```

```{r}
DimPlot(seurat, group.by="CellType", label=T)
DimPlot(seurat, group.by="seurat_clusters", label=T)
```

```{r}
breast_bassez_cohort1 <- getEpithelial(seurat)
```

## Cohort 2

```{r}
seurat <- readRDS("../output/tumour_annotations/breast_bassez_cohort2_seurat.rds")
#Add some metadata to help post-merge
seurat$Cancer <-  "Breast"
seurat$Source <- "Bassez"
```

```{r}
DimPlot(seurat, group.by="CellType", label=T)
DimPlot(seurat, group.by="seurat_clusters", label=T)
```

```{r}
breast_bassez_cohort2 <- getEpithelial(seurat)
```

# SCC - Ji et al - GSE144236

```{r}
seurat <- readRDS("../output/tumour_annotations/scc_ji_seurat.rds")
#Add some metadata to help post-merge
seurat$Cancer <-  "SCC"
seurat$Source <- "Ji"
```

```{r}
DimPlot(seurat, group.by="CellType", label=T)
DimPlot(seurat, group.by="seurat_clusters", label=T)
```

```{r}
scc_ji <- getEpithelial(seurat)
```

# Nasopharyngeal Carcinoma - Chen et al - GSE150430

```{r}
seurat <- readRDS("../output/tumour_annotations/npc_chen_seurat.rds")
#Add some metadata to help post-merge
seurat$Cancer <-  "NPC"
seurat$Source <- "Chen"
```

```{r}
DimPlot(seurat, group.by="CellType", label=T)
DimPlot(seurat, group.by="seurat_clusters", label=T)
```

```{r}
npc_chen <- getEpithelial(seurat)
```


# Combine all
```{r}
seurat <- merge(colorectal_lee, list(colorectal_lee_kul, colorectal_uhlitz,
                              colorectal_qian, gastric_sathe, lung_kim, lung_lambrechts,
                              lung_laughney, lung_qian, lung_wu, ovarian_geistlinger,
                              ovarian_qian, pdac_steele, breast_wu, breast_qian,
                              breast_bassez_cohort1, breast_bassez_cohort2, scc_ji,
                              npc_chen),
                merge.data=F)
seurat$SampleID <- paste(seurat$Cancer, seurat$Source, seurat$Patient, sep="_")


rm(colorectal_lee, colorectal_lee_kul, colorectal_uhlitz,
                              colorectal_qian, gastric_sathe, lung_kim, lung_lambrechts,
                              lung_laughney, lung_qian, lung_wu, ovarian_geistlinger,
                              ovarian_qian, pdac_steele, breast_wu, breast_qian,
                              breast_bassez_cohort1, breast_bassez_cohort2, scc_ji,
                              npc_chen)


samples_keep <- names(table(seurat$SampleID)[which(table(seurat$SampleID) > 100)])

print(paste0("Keeping ", length(samples_keep), "/", length(unique(seurat$SampleID)), " samples"))

seurat <- subset(seurat, subset = SampleID %in% samples_keep)


seurat <- NormalizeData(seurat)
seurat <- FindVariableFeatures(seurat)
seurat <- ScaleData(seurat, vars.to.regress=c("S.Score", "G2M.Score", "percent.mito",
                                              "nCount_RNA"))
seurat <- RunPCA(seurat, verbose=F)
seurat <- RunUMAP(seurat, dims=1:30)

saveRDS(seurat, file="../output/combined_epi_revised.rds")
```

```{r}
DimPlot(seurat, group.by="Cancer", pt.size=0.1)
```








