---
title: "EMP Regulation"
output: html_notebook
---

# Goal 
Will do pathway and TF regulon inference on EMP programs from tumour scRNA-seq data

# Dependencies
```{r}
library(Seurat)
library(progeny)
library(viper)
library(dorothea)
library(tidyverse)
library(factoextra)
```

# Load the data
```{r}
seurat <- readRDS("../data/pan_cancer_epi.rds")
sample_list <- unique(seurat$SampleID)
meta_list <- readRDS("~/Projects/emt_programs/output/master_metadata.rds")
emp_programs <- readLines("../output/emp_program_list.txt")
emt_sig <- readLines("../output/conserved_emt_signature.txt")
```

```{r}
# Useful color palettes
cols <- paletteer::paletteer_c("scico::berlin", n=100)
```


# PROGENy of programs
Goal is to find pathway activity scores that correlate with program footprints

I'm not positive it matters, but I'm going to run it separately on each tumour sample because it scales the result and I don't want anything weird between samples affecting it

We'll run PROGENy on the pearson residuals of the SCT model for each sample
```{r}
for(i in 1:length(sample_list)){
  seurat_subset <- subset(seurat, subset = SampleID == sample_list[i])
  seurat_subset <- SCTransform(seurat_subset, vars.to.regress="percent.mito")
  progeny_mat <- progeny(as.matrix(seurat_subset[["SCT"]]@scale.data),
                         top=500)
  meta_list[[i]] <- cbind(meta_list[[i]], progeny_mat)
}
```

```{r}
saveRDS(meta_list, file="../output/master_metadata.rds")
```

## Activity ~ Program
Because each program has a different
```{r}
#pathways <- colnames(progeny_mat)
pathways <- c("Androgen", "EGFR", "Estrogen", "Hypoxia", "JAK-STAT", "MAPK", "NFkB", "p53", "PI3K", "TGFb", "TNFa", "Trail", "VEGF", "WNT")

testPathway <- function(pathway, program, df){
  score_norm <- df[,program] / max(df[,program])
  test_summary <- summary(lm(df[,pathway] ~ score_norm))
  test_res <- data.frame(Coef = test_summary$coefficients[2,1],
                         pval = test_summary$coefficients[2,4],
                         Pathway = pathway)
  return(test_res)
}

runPathwayTests <- function(sampleID){
  print(paste0("Testing: ", sampleID))
  df <- meta_list[[sampleID]]
  total_arch <- unique(paste0("A", unique(df$Archetype_Membership)))
  total_arch <- total_arch[order(total_arch)]
  
  pathway_test_res <- list()
  for(i in 1:length(total_arch)){
    test_summary <- lapply(pathways, testPathway, program = total_arch[i], df = df)
    test_summary <- do.call("rbind", test_summary)
    test_summary$padj <- p.adjust(test_summary$pval, method="BH")
    test_summary$Program <- paste0(sampleID, "_", total_arch[i])
    pathway_test_res[[i]] <- test_summary
  }
  
  pathway_test_res <- do.call("rbind", pathway_test_res)
}
```

```{r}
pathway_scores <- lapply(sample_list, runPathwayTests)
pathway_scores <- do.call("rbind", pathway_scores)
```

### Coef matrix

```{r}
pathway_coef <- pathway_scores
pathway_coef$padj <- NULL
pathway_coef$pval <- NULL
pathway_coef <- pivot_wider(pathway_coef, names_from=Pathway, values_from=Coef)
pathway_coef <- as.data.frame(pathway_coef)
rownames(pathway_coef) <- pathway_coef$Program
pathway_coef$Program <- NULL
pathway_coef <- as.matrix(pathway_coef)
```

```{r}
#If you have the emt gene set correlation matrix and want to order it the same:
#heatmap_order <- rownames(cor_list_merge)[emt_cor_heatmap$tree_row$order]
```

```{r}
#Skip row clustering if you want to use the ordering from the EMT correlation heatmap
pathway_heatmap <- pheatmap::pheatmap(pathway_coef,
                                      color = cols,
                                      breaks=seq(-3, 3, length.out=101),
                                      cluster_rows=T,
                                      cluster_cols=T,
                                      show_rownames=F,
                                      cutree_row=4,
                                      fontsize = 12,
                                      angle_col=45,
                                      clustering_method="ward.D2",
                                      filename="../figs/program_pathway_activity_clustered_heatmap.png",
                                      width=5, height=8)

plot(pathway_heatmap$gtable)
```

## PCA of coefficients
```{r}
pca_res <- prcomp(pathway_coef, scale=F)
```

```{r}
fviz_eig(pca_res) + ylab("Percentage variance explained") + 
  theme(axis.text = element_text(size=10, color="black"),
        axis.title = element_text(size=12))
```

Just to get a factor corresponding to whether the programs are EMP programs or not

```{r}
emt_program <- factor(rownames(pathway_coef) %in% emp_programs)
```


```{r}
pca1 <- fviz_pca_biplot(pca_res, geom.ind="point", axes=c(1,2),
                col.ind = emt_program, col.var = "black",
                repel=T) +
  scale_color_manual(values=c('lightgrey', 'firebrick'))

pca2 <- fviz_pca_biplot(pca_res, geom.ind="point", axes=c(1,3),
                col.ind = emt_program, col.var = "black",
                repel=T) +
  scale_color_manual(values=c('lightgrey', 'firebrick'))
pca1
pca2
```

```{r}
ggsave(pca1, filename="../figs/progeny_program_coef_pca_1_2.pdf",
       device = cairo_pdf(),
       width=5, height=3.5)
ggsave(pca2, filename="../figs/progeny_program_coef_pca_1_3.pdf",
       device = cairo_pdf(),
       width=5, height=3.5)
```

# DoRothEA
Still not sure if I'm going to be happier with DoRothEA or SCENIC/AUCell. DoRothEA is super fast though, so no harm.

```{r}
data(dorothea_hs, package="dorothea")
```

```{r}
regulons <- dorothea_hs %>%
  filter(confidence %in% c("A", "B", "C"))
```

## Score each sample
```{r}
runRegulon <- function(sample_name){
  df <- meta_list[[sample_name]]
  cells <- rownames(df)
  
  exp_mat <- as.matrix(seurat[["RNA"]]@data[,cells])
  regulon_scores <- run_viper(exp_mat, regulons,
                              tidy=F,
                              options = list(minsize=5,
                                             cores=2,
                                             verbose=T))
  regulon_scores <- t(regulon_scores)
}
```

TF scores are a normalized enrichment score, so there should be value in keeping them as this rather than Z-scoring them.

```{r}
regulon_score_list <- lapply(sample_list, runRegulon)
names(regulon_score_list) <- sample_list
saveRDS(regulon_score_list, file="../output/regulon_cell_scores.rds")
```

## Differential activity
```{r}
testRegulon <- function(tf_name, arch_score, regulon_mat){
  regulon_score <- regulon_mat[,tf_name]
  model_res <- summary(lm(regulon_score ~ arch_score))
  res_table <- data.frame(pval = model_res$coefficients[2,4],
                          coef = model_res$coefficients[2,1],
                          TF = tf_name)
  return(res_table)
}

testAllRegulons <- function(sampleID){
  print(paste0("Testing: ", sampleID))
  
  # Get list of archetypes for sample
  meta <- meta_list[[sampleID]]
  arch_list <- unique(paste0("A", meta$Archetype_Membership))
  
  #Get TF list
  tf_list <- colnames(regulon_score_list[[sampleID]])
  
  print("Testing for TFs associated with archetype program activity")
  result_table <- list()
  for(i in 1:length(arch_list)){
    arch_score <- as.numeric(meta[,arch_list[i]])
    #Normalize archetype scores to have max = 1
    arch_score <- arch_score / max(arch_score)
    arch_res <- lapply(tf_list, testRegulon, arch_score = arch_score, regulon_mat = regulon_score_list[[sampleID]])
    arch_res <- do.call("rbind", arch_res)
    arch_res$padj <- p.adjust(arch_res$pval, method="BH")
    arch_res$Program <- paste0(sampleID, "_", arch_list[i])
    result_table[[i]] <- arch_res
  }
  result_table <- do.call("rbind", result_table)
}
```

```{r}
regulon_test_res <- lapply(sample_list, testAllRegulons)

regulon_test_res <- do.call("rbind", regulon_test_res)
write.csv(regulon_test_res, file="../output/regulon_differential_activity.csv", row.names = F)
```

## EMP TFs
Just a quick look
```{r}
regulon_test_res %>% 
  filter(padj <= 0.05 & Program %in% emp_programs & coef > 0.5) %>%
  group_by(TF) %>% summarise(count = n()) %>% arrange(desc(count))
```



## Convert to matrix
```{r}
regulon_coef <- regulon_test_res
regulon_coef$padj <- NULL
regulon_coef$pval <- NULL
regulon_coef <- pivot_wider(regulon_coef, names_from=TF, values_from=coef)
regulon_coef <- as.data.frame(regulon_coef)
rownames(regulon_coef) <- regulon_coef$Program
regulon_coef$Program <- NULL
regulon_coef <- as.matrix(regulon_coef)
```

```{r}
annotation_row <- data.frame(EMP_Program = ifelse(rownames(regulon_coef) %in% emp_programs, "EMP", "Other"))
rownames(annotation_row) <- rownames(regulon_coef)
pheatmap(regulon_coef[emp_programs,sig_tfs],
         color=cols,
        # annotation_row = annotation_row,
         breaks = seq(-2, 2, length.out=101),
         cluster_rows=T,
         cluster_cols=T,
         clustering_method="ward.D2",
         show_rownames=F,
         show_colnames=T,
         angle_col = 45,
         fontsize_col = 4,
         filename="~/Downloads/tf_heatmap.png",
         width=20, height=6)
```


Only keep TFs sig in EMP programs
```{r}
sig_tfs <- regulon_test_res %>% filter(padj <= 0.05 &
                                         Program %in% emp_programs &
                                         coef > 1) %>%
  group_by(TF) %>% 
  summarise(count = n()) %>%
  arrange(desc(count)) %>%
  filter(count > 10) %>%
  pull(TF)
```


## PCA of coefficients
```{r}
sig_tfs <- regulon_test_res %>% filter(padj <= 0.05 &
                                         Program %in% emp_programs &
                                         abs(coef) > 1) %>%
  group_by(TF) %>% 
  summarise(count = n()) %>%
  arrange(desc(count)) %>%
  #filter(count > 10) %>%
  pull(TF)
```


```{r}
pca_res <- prcomp(regulon_coef[,sig_tfs], scale=F)
```

```{r}
fviz_eig(pca_res) + ylab("Percentage variance explained") + 
  theme(axis.text = element_text(size=10, color="black"),
        axis.title = element_text(size=12))
```

Just to get a factor corresponding to whether the programs are EMP programs or not

```{r}
emt_program <- factor(rownames(regulon_coef) %in% emp_programs)
```


Get TFs to label

```{r}
tmp <- pca_res$rotation[,1:2]
ggplot(as.data.frame(tmp), aes(PC1, PC2)) +
  geom_point(color = "black") +
  geom_vline(xintercept=-0.025) +
  geom_hline(yintercept = 0.025) +
  theme_bw()
sig_tfs <- rownames(tmp)[tmp[,1] <= -0.025 & tmp[,2] >= 0.025]
```


```{r}
#OR
var_dat <- get_pca_var(pca_res)
tmp <- var_dat$coord[,1:2]
#TFs with strongest "up and to the left" loading
sig_tfs <- rownames(tmp)[tmp[,1] <= -0.025 & tmp[,2] >= 0.25] 
sig_tfs
```

```{r}
tf_labels = c("KLF5", "TEAD1", "KLF6", "BCL3", "HIF1A",
              "MEF2A", "JUNB", "STAT2", "TEAD4", "TCF12",
              "STAT3", "RELA", "ESR1",
              "SMAD2", "SOX13", "TWIST1", "TP53")
```


```{r}
pca1 <- fviz_pca_biplot(pca_res, geom.ind="point", axes=c(1,3),
                col.ind = emt_program, repel=T,
               # select.var = list(name=tf_labels),
                labelsize=2) +
  scale_color_manual(values=c('lightgrey', 'firebrick'))
```

```{r}
ggsave(pca1, filename="../figs/dorothea_pca_huge.pdf",
       device=cairo_pdf(), width=10, height=10)
ggsave(pca1, filename="../figs/dorothea_pca.pdf",
       device=cairo_pdf(), width=5, height=3.5)
```



