---
title: "Pan cancer EMT program"
output: html_notebook
---

```{r}
library(Seurat)
library(tidyverse)
library(ACTIONet)
library(pheatmap)
library(ComplexHeatmap)
```

# Load the data
Objects generated in actionet_pan_cancer.Rmd

```{r}
seurat <- readRDS("../data/pan_cancer_epi.rds")
sample_list <- unique(seurat$SampleID)
meta_list <- readRDS("../output/master_metadata.rds")
```

# Archetype-associated genes
I had originally used the specificity scores computed by ACTIONet, but I'm feeling less comfortable because I really don't know how they're calculated. To get a little bit more control, the plan is to go archetype by archetype, running a linear model on the SCT pearson residuals of the top 2k genes to find genes significantly associated with the archetype footprint scores. Then I can use both p-values and the coefficients to narrow in on things.

I want to keep mitochondrial genes out of the analysis for convenience
```{r}
mtGenes <- grep('^MT-',rownames(seurat),value=TRUE)
```

# Functions
```{r}
##### Linear model. Can change to test any other test
testGene <- function(gene, arch_score, mito, seu){
  model_res <- summary(lm(seu[["RNA"]]@data[gene,] ~ arch_score + mito))
  res_table <- data.frame(pval = model_res$coefficients[2,4],
                          coef = model_res$coefficients[2,1],
                          Gene = gene)
  return(res_table)
}

##### Big function to get gene scores for each program of a given sample
testPrograms <- function(sampleID){
  print(paste0("Testing: ", sampleID))
  
  #Subset seurat
  print("Getting variable features")
  seurat_subset <- subset(DietSeurat(seurat), subset = SampleID == sampleID)
  seurat_subset <- FindVariableFeatures(seurat_subset)
  var_dat <- seurat_subset[["RNA"]]@meta.features
  var_dat$pct <- rowSums(seurat_subset[["RNA"]]@data > 0) / ncol(seurat_subset)
  var_dat <- var_dat %>%
    filter(pct > 0.05) %>% #detection >5%
    arrange(desc(vst.variance.standardized))
  top_genes <- rownames(var_dat)[1:2000]
  
  # Get archetype data
  meta <- meta_list[[sampleID]]
  arch_list <- unique(paste0("A", meta$Archetype_Membership))
  
  # Test genes for each archetype. 
  print("Testing for genes associated with archetype program activity")
  result_table <- list()
  for(i in 1:length(arch_list)){
    arch_score <- as.numeric(meta[,arch_list[i]])
    mito <- as.numeric(meta$percent.mito)
    #Normalize archetype scores to have max = 1
    arch_score <- arch_score / max(arch_score)
    arch_res <- lapply(top_genes, testGene, 
                       arch_score = arch_score,
                       mito = mito, 
                       seu = seurat_subset)
    arch_res <- do.call("rbind", arch_res)
    arch_res$padj <- p.adjust(arch_res$pval, method="BH")
    arch_res$Program <- paste0(sampleID, "_", arch_list[i])
    result_table[[i]] <- arch_res
  }
  result_table <- do.call("rbind", result_table)
}
```

```{r}
program_scores <- lapply(sample_list, testPrograms)

program_scores <- do.call("rbind", program_scores)
write.csv(program_scores, file="../output/archetype_program_scores.csv", row.names = F)
```

```{r}
program_scores <- read.csv("../output/archetype_program_scores.csv")
```


# Organize program scores
## -log10(padj) mat
For this matrix, I think I'll only keep significant changes w/ positive coefficients so that we don't get genes with hits in multiple programs without knowing that they're going in opposite directions.

```{r}
program_pval <- program_scores
program_pval <- filter(program_pval, coef > 0 & padj <= 0.05) # torn about doing only sig--should help
program_pval$coef <- NULL
program_pval$pval <- NULL
program_pval$padj <- -log10(program_pval$padj)
program_pval <- pivot_wider(program_pval, names_from=Program, values_from=padj)
program_pval <- as.data.frame(program_pval)
rownames(program_pval) <- program_pval$Gene
program_pval$Gene <- NULL
program_pval <- as.matrix(program_pval)
program_pval[is.na(program_pval)] <- 0
```

## Coefficient matrix
For this, I'll keep positive and negative values for the purpose of correlating programs

```{r}
program_coef <- program_scores
program_coef$padj <- NULL
program_coef$pval <- NULL
program_coef <- pivot_wider(program_coef, names_from=Program, values_from=coef)
program_coef <- as.data.frame(program_coef)
rownames(program_coef) <- program_coef$Gene
program_coef$Gene <- NULL
program_coef <- as.matrix(program_coef)
program_coef[is.na(program_coef)] <- 0
```

## List of program-associated genes
```{r}
assessSig <- function(program){
  hits <- program_scores %>%
    filter(Program == program) %>%
    filter(padj <= 0.05 & coef > 0.5) %>%
    arrange(desc(padj)) %>% #Just so genes are actually ranked by significance in the list
    pull(Gene)#can add coefficient cutoffs if desired
  return(hits)
}

program_genes <- lapply(colnames(program_pval), assessSig)
names(program_genes) <- colnames(program_pval)
```

# Program size (sig count)
```{r}
program_size <- program_scores %>%
  filter(padj <= 0.05 & coef > 0.5) %>%
  group_by(Program) %>%
  summarise(Size = length(Gene))
```

```{r}
hist(program_size$Size, breaks=100)
```



# Gene counts
```{r}
gene_counts <- program_scores %>%
  filter(padj <= 0.05 & coef > 0.5) %>%
  group_by(Gene) %>%
  summarise(Count = length(Program)) %>%
  arrange(desc(Count))

gene_counts
```


# Correlate programs
I could just take programs_coef and correlate, but I feel like this would diminish some of the signal given that the matrix has ~15k genes and only ~2k are tested per program. The many 0,0s could mess with it. I think I could try two options:
1) Only correlate on genes tested for the pairwise conditions (max 4k genes)
2) Only correlate on genes significant in at least one of the two conditions


```{r}
program_list <- unique(program_scores$Program)
```


## 1) Only tested genes
It's really slow because I have to calculate all pairwise separately and can't just feed a full matrix
I think it's the dplyr function that adds up here
```{r}
corTest <- function(program1, program2){
  genes_test <- filter(program_scores, Program %in% c(program1, program2)) %>%
    pull(Gene) %>%
    unique()
  dat_mat <- program_coef[genes_test, c(program1, program2)]
  program_cor <- cor(dat_mat[,program1], dat_mat[,program2], method="spearman")
  return(program_cor)
}

testCorrelation <- function(program){
  cor_vals <- lapply(program_list, corTest, program2 = program)
  cor_vals <- do.call("rbind", cor_vals)
  rownames(cor_vals) <- program_list
  return(cor_vals)
}

cor_mat <- lapply(program_list, testCorrelation)
cor_mat <- do.call("cbind", cor_mat)
colnames(cor_mat) <- program_list


write.csv(cor_mat, file="../output/program_coef_correlation.csv")
```

```{r}
cols <- paletteer::paletteer_c("scico::berlin", n=100)
pheatmap(cor_mat,
         color = cols,
         breaks = seq(-0.5, 0.5, length.out=101),
         clustering_method="ward.D2",
         show_colnames = F,
         show_rownames = F,
         filename="../figs/archetype_program_correlation.png",
         width=8, height=8) 
```


## Jaccard
```{r}
getJaccard <- function(program){
  jaccard_scores <- 0
  for(i in 1:length(program_genes)){
    jaccard_scores[i] <- length(intersect(program_genes[[program]], program_genes[[i]])) / length(union(program_genes[[program]], program_genes[[i]]))
  }
  names(jaccard_scores) <- names(program_genes)
  return(jaccard_scores)
}
```

```{r}
jaccard <- lapply(names(program_genes), getJaccard)
jaccard <- do.call(rbind, jaccard)
rownames(jaccard) <- colnames(jaccard)
```

```{r}
pheatmap::pheatmap(jaccard,
         color = viridis::inferno(100),
         border_color = "black",
         breaks = seq(0, 0.50, length.out=101),
         clustering_method="ward.D2",
         show_colnames = T,
         show_rownames = F,
         angle_col = 45,
         fontsize_col = 4,
         filename="../figs/archetype_program_jaccard_emp_programs.png",
         width=12, height=12) 
```

# Get EMT programs
```{r}
cor_list_merge <- read.csv("../output/archetype_emt_correlation.csv", row.names=1)
cor_list_merge <- as.matrix(cor_list_merge)
```

```{r}
#cols <- paletteer::paletteer_c("scico::berlin", n=100)
cols <- colorRampPalette(rev(RColorBrewer::brewer.pal(9, "RdBu")))(100)
```

```{r}
gene_set_order <- c("Taube", "Tan_EMT", "GO", "CancerSEA", "Kinker_II", "Kinker_III",
                    "Hallmark", "Puram_pEMT", "dbEMT", "Cook", "Kinker_I")
```

```{r}
emt_cor_heatmap <- pheatmap::pheatmap(cor_list_merge[,gene_set_order],
                                      color = cols,
                                      breaks=seq(-0.75, 0.75, length.out=101),
                                      cluster_rows=T,
                                      cluster_cols=F,
                                      show_rownames=F,
                                      cutree_row=4,
                                      fontsize = 12,
                                      angle_col=45,
                                      clustering_method="ward.D2",
                                      #filename="../figs/archetype_emt_correlation_all.png",
                                      width=3.5, height=8)

plot(emt_cor_heatmap$gtable)
```

```{r}
clusters <- as.data.frame(cutree(emt_cor_heatmap$tree_row, k=4))
colnames(clusters) <- "Cluster"
clusters$Program <- rownames(clusters)
table(clusters$Cluster)
```

Order from top to bottom
```{r}
unique(clusters$Cluster[emt_cor_heatmap$tree_row$order])
```

```{r}
emt_arch_high <- clusters %>% filter(Cluster %in% c(4)) %>% pull(Program)
emt_arch_broad <- clusters %>% filter(Cluster %in% c(3,4)) %>% pull(Program)
```

```{r}
writeLines(emt_arch_high, file("../output/emp_program_list_highConf.txt"))
writeLines(emt_arch_broad, file("../output/emp_program_list_broad.txt"))
```


# Jaccard of EMP programs
```{r}
emp_program_genes <- program_genes[emt_arch_broad]
```

```{r}
getJaccard <- function(program){
  jaccard_scores <- 0
  for(i in 1:length(emp_program_genes)){
    jaccard_scores[i] <- length(intersect(emp_program_genes[[program]], emp_program_genes[[i]])) / length(union(emp_program_genes[[program]], emp_program_genes[[i]]))
  }
  names(jaccard_scores) <- names(emp_program_genes)
  return(jaccard_scores)
}
```

```{r}
jaccard <- lapply(names(emp_program_genes), getJaccard)
jaccard <- do.call(rbind, jaccard)
rownames(jaccard) <- colnames(jaccard)
```

```{r}
pheatmap::pheatmap(jaccard,
         color = viridis::inferno(100),
         border_color = "black",
         breaks = seq(0, 0.3, length.out=101),
         clustering_method="ward.D2",
         show_colnames = F,
         show_rownames = T,
         fontsize_col = 3,
         filename="../figs/archetype_program_jaccard_emp_programs.png",
         width=15, height=15) 
```

# Coefs of genes associated with EMT programs

Goal here: defining a signature of the strongest EMT-associated genes. For this, we only really care to consider the genes with pretty strong coefficients in the model. 

```{r}
#Get relevant genes
genes_keep <- program_scores %>%
  filter(Program %in% emt_arch_broad &
           padj <= 0.05 &
           abs(coef) > 2) %>%
  pull(Gene) %>%
  unique()

emt_scores <- program_coef[genes_keep, emt_arch_broad]
#emt_scores <- program_coef[emt_sig, emt_arch_broad]
emt_scores[emt_scores < -3] <- -3
emt_scores[emt_scores > 3] <- 3
```

```{r}
cols <- colorRampPalette(rev(RColorBrewer::brewer.pal(9, "RdBu")))(100)
emt_pval <- pheatmap::pheatmap(emt_scores,
                     color = cols,
                     breaks=seq(-3, 3, length.out=101),
                     clustering_method="ward.D2",
                     show_colnames=T,
                     angle_col = 45,
                     fontsize = 2.5,
                     show_rownames=F,
                     cutree_rows=2,
                     filename="../figs/emt_program_coefficients.png",
                     width=15, height=10)
```

```{r}
emt_clusters <- as.data.frame(cutree(emt_pval$tree_row, k=2))
colnames(emt_clusters) <- "Cluster"
emt_clusters$Gene <- rownames(emt_clusters)
table(emt_clusters$Cluster)
```

Order from top to bottom
```{r}
unique(emt_clusters$Cluster[emt_pval$tree_row$order])
```

```{r}
emt_sig <- emt_clusters %>% filter(Cluster==2) %>% pull(Gene)
emt_sig <- emt_sig[-grep("^MT-", emt_sig)]
writeLines(emt_sig, file("../output/conserved_emt_signature.txt"))
emt_sig
```

This coefficient approach is probably the best for designing a signature (especially if goal is to use as a scRNA-seq signature). Will contain the genes with strongest signal-to-noise, but doesn't necessarily mean that these are the only genes involved. Eg. Lowering coefficient threshold allows some TFs to be significant more often. Plus, a 233 gene set is a good number for scoring.

Interesting to see that MHC class I genes are consistently there. Actually, I thought MHC-I genes are often downregulated. Hm.

# Signature by frequency
```{r}
gene_counts <- program_scores %>%
  filter(Program %in% emt_arch_broad & padj <= 0.05 & coef > 1) %>%
  group_by(Gene) %>%
  summarise(Count = n(), avg_coef = mean(coef)) %>% 
  arrange(desc(Count))
gene_counts$Rank <- 1:nrow(gene_counts)
```

```{r}
gene_counts
```

```{r}
plot(gene_counts$Count)
abline(h=20)
```


```{r}
emt_sig <- gene_counts %>%
  filter(Count >= 20) %>%
  pull(Gene)
emt_sig <- emt_sig[-grep("^MT-", emt_sig)]
length(emt_sig)
```


# Refine With Count
The idea here is that some of these genes go down in a high proportion of programs, or only go up in a small number. Can we refine the signature by selecting those that go up in a reasonable number of programs and do not go down in many

```{r}
gene_counts_up <- program_scores %>%
  filter(Program %in% emt_arch_broad & Gene %in% emt_sig) %>%
  filter(padj <= 0.05 & coef > 1) %>%
  group_by(Gene) %>%
  summarise(Count_Up = length(Program)) %>%
  arrange(desc(Count_Up))


gene_counts_down<- program_scores %>%
  filter(Program %in% emt_arch_broad  & Gene %in% emt_sig) %>%
  filter(padj <= 0.05 & coef < -1) %>%
  group_by(Gene) %>%
  summarise(Count_Down = length(Program)) %>%
  arrange(desc(Count_Down))

gene_counts <- left_join(gene_counts_up, gene_counts_down,
                         by="Gene")
gene_counts$Count_Down[is.na(gene_counts$Count_Down)] <- 0
```


```{r}
ggplot(gene_counts, aes(Count_Up, Count_Down)) +
  geom_point(alpha=0.5,) +
  geom_hline(yintercept = 10, linetype=2) +
  #geom_vline(xintercept = 10, linetype=2) +
  theme_classic()
```


```{r}
labels <- c("SOX4", "IL32", "TGFBI", "")
```

```{r}
gene_counts$Label <- ""
gene_counts$Label[which(gene_counts$Gene %in% labels)] <- gene_counts$Gene[which(gene_counts$Gene %in% labels)]
gene_counts <- filter(gene_counts, Gene %in% emt_sig) # Only plot those that have already passed EMT sig

ligand_freq_plot <- ggplot(gene_counts, aes(Count_Up, Count_Down, label=Label)) +
  geom_point(size=1, color="black") +
  geom_text_repel(colour="firebrick",
                  max.overlaps = Inf,
                  min.segment.length=0.2,
                  direction="both",
                  nudge_y=4,
                  force=8,
                  size=3.5,
                  hjust=0.5,
                  segment.size=0.5,
                  segment.color="firebrick") +
  geom_hline(yintercept = 10, linetype=2, color="black") +
  geom_vline(xintercept = 10, linetype=2, color="black") +
  xlab("EMP Upregulated") + ylab("EMP Downregulated") +
  #scale_y_continuous(breaks=c(0,20, 40, 60, 80), limits=c(0,88)) +
  theme_classic() +
  theme(axis.text = element_text(size=10, color="black"),
        axis.title = element_text(size=12))

#ggsave(ligand_freq_plot, filename="../figs/emp_gene_counts.pdf",
#       width=2.7, height=3.2, device=cairo_pdf())
```

```{r}
ligand_freq_plot
```

#Refine EMT signature
```{r}
emt_sig <- gene_counts %>%
  filter(Count_Down <= 10) %>%
  pull(Gene)

writeLines(emt_sig, file("../output/conserved_emt_signature.txt"))
```

# Coef Heatmap of Sig
```{r}
emt_scores <- program_coef[emt_sig, emt_arch_broad]
#emt_scores <- program_coef[emt_sig, emt_arch_broad]
emt_scores[emt_scores < -3] <- -3
emt_scores[emt_scores > 3] <- 3
```

```{r}
cols <- colorRampPalette(rev(RColorBrewer::brewer.pal(9, "RdBu")))(100)
emt_pval <- pheatmap::pheatmap(emt_scores,
                     color = cols,
                     breaks=seq(-3, 3, length.out=101),
                     clustering_method="ward.D2",
                     show_colnames=T,
                     angle_col = 45,
                     fontsize = 2.5,
                     show_rownames=F,
                     cutree_rows=2,
                     filename="../figs/emt_sig_program_coefficients.png",
                     width=15, height=10)
```


# Total genes per sample
```{r}
assessSig <- function(program){
  hits <- program_scores %>%
    filter(Program == program) %>%
    filter(padj <= 0.05 & coef > 1) %>%
    arrange(desc(padj)) %>% #Just so genes are actually ranked by significance in the list
    pull(Gene)#can add coefficient cutoffs if desired
  return(hits)
}

program_list <- unique(program_scores$Program)
program_genes <- lapply(program_list, assessSig)
names(program_genes) <- program_list
```

```{r}
emt_program_names <- readLines("~/Projects/emt_programs/output/emp_program_list.txt")
emt_program <- program_genes[emt_program_names]
```

```{r}
dat <- unlist(lapply(emt_program, length))
dat <- data.frame(Program = names(dat), Count = dat)
dat$Program <- substr(dat$Program, 1, nchar(dat$Program)-3)
dat <- dat %>% group_by(Program) %>% summarise(total = sum(Count))
dat
```


# GO Enrichment of EMT Signature
```{r}
go <- fgsea::gmtPathways("~/Data/GeneLists/GOTerms.BP.v6.1.symbols.gmt")
```

```{r}
library(GeneOverlap)
enrichmentTest <- function(gene_set){
  go_object <- newGeneOverlap(emt_sig, gene_set,
                              spec = "hg19.gene")
  go_object <- testGeneOverlap(go_object)
  results <- go_object@pval
}

enrichment <- lapply(go, enrichmentTest)
enrichment <- as.data.frame(do.call("rbind", enrichment))
colnames(enrichment) <- "pval"
enrichment$padj <- p.adjust(enrichment$pval, method = "BH")
```

```{r}
table(enrichment$padj <= 0.05)
```

```{r}
enrichment$Term <- rownames(enrichment)
```

```{r}
enrichment$Term <- gsub("GO_", "", enrichment$Term)
enrichment$Term <- gsub("_", " ", enrichment$Term)
enrichment$Term <- Hmisc::capitalize(tolower(enrichment$Term))
```

```{r}
enrichment %>% arrange(padj) %>% head(300)
```

```{r}
plot_terms <- c("Regulation of cell death", "Regulation of cell proliferation",
                "Epithelium development", "Regulation of cellular component movement",
                "Regulation of cell adhesion", "Immune system process", "Cell motility",
                "Response to wounding", "Vasculature development")
plot_dat <- filter(enrichment, Term %in% plot_terms) %>% arrange(padj)
plot_dat$Term <- factor(plot_dat$Term, levels=plot_dat$Term)

enrichment_plot <- ggplot(plot_dat, aes(x=-log10(padj), y=Term)) +
  geom_col(width=0.025, color="grey20", fill="grey20") +
  geom_point(size=5, shape=21, color="black", fill="firebrick") +
  geom_vline(xintercept=-log10(0.05), linetype=2) +
  scale_x_continuous(expand=c(0,0), limits=c(0, max(-log10(plot_dat$padj)+0.85))) +
  xlab("-log10(p-value)") + ylab("") +
  theme_classic() +
  theme(axis.text=element_text(size=12, color="black"),
        axis.title=element_text(size=12),
        legend.position="none")
```

```{r}
enrichment_plot
```

```{r}
ggsave(enrichment_plot, device=cairo_pdf(),
       filename="../figs/emt_sig_go.pdf",
       width=6.5, height=3)
```

# GO Enrichment of EMT DOWN Signature

```{r}
enrichmentTest <- function(gene_set){
  go_object <- newGeneOverlap(emt_down_sig, gene_set,
                              spec = "hg19.gene")
  go_object <- testGeneOverlap(go_object)
  results <- go_object@pval
}

enrichment <- lapply(go, enrichmentTest)
enrichment <- as.data.frame(do.call("rbind", enrichment))
colnames(enrichment) <- "pval"
enrichment$padj <- p.adjust(enrichment$pval, method = "BH")
```

```{r}
table(enrichment$padj <= 0.05)
```

```{r}
enrichment$Term <- rownames(enrichment)
```

```{r}
enrichment$Term <- gsub("GO_", "", enrichment$Term)
enrichment$Term <- gsub("_", " ", enrichment$Term)
enrichment$Term <- Hmisc::capitalize(tolower(enrichment$Term))
```

```{r}
enrichment %>% arrange(padj) %>% head(200)
```

```{r}
plot_terms <- c("Cell cycle", "Proteasomal protein catabolic process",
                "Protein targeting", "Amide biosynthetic process", "Cellular catabolic process", 
                "Membrane organization")
plot_dat <- filter(enrichment, Term %in% plot_terms) %>% arrange(padj)
plot_dat$Term <- factor(plot_dat$Term, levels=plot_dat$Term)

enrichment_plot <- ggplot(plot_dat, aes(x=-log10(padj), y=Term)) +
  geom_col(width=0.025, color="grey20", fill="grey20") +
  geom_point(size=5, shape=21, color="black", fill="darkblue") +
  geom_vline(xintercept=-log10(0.05), linetype=2) +
  scale_x_continuous(expand=c(0,0), limits=c(0, max(-log10(plot_dat$padj)+2))) +
  xlab("-log10(p-value)") + ylab("") +
  theme_classic() +
  theme(axis.text=element_text(size=12, color="black"),
        axis.title=element_text(size=12),
        legend.position="none")
```

```{r}
enrichment_plot
```
```{r}
ggsave(enrichment_plot, device=cairo_pdf(),
       filename="../figs/emt_sig_downGenes_go.pdf",
       width=5, height=2)
```

# Score cells with signatures
```{r}
seurat <- AddModuleScore_UCell(seurat, features=list(EMT_Sig = emt_sig),
                                  name="")
saveRDS(seurat, file="../data/pan_cancer_epi.rds")
```




## FeaturePlot
```{r}
dat <- seurat@meta.data
dat$UMAP1 <- Embeddings(seurat, 'umap')[,1]
dat$UMAP2 <- Embeddings(seurat, 'umap')[,2]
dat$EMT_Sig1[dat$EMT_Sig1 < 0] <- 0
dat$EMT_Sig1[dat$EMT_Sig1 > 1] <- 1
```

```{r}
emt_sig_umap <- ggplot(dat, aes(UMAP1, UMAP2)) +
  geom_point(shape=16, size=0.1, aes(colour=EMT_Sig1), alpha=0.5) +
  scale_colour_gradientn(colours=c(viridis::inferno(100))) +
  #scale_color_gradientn(colours=c('lightgrey', 'red')) +
  theme_void() +
  theme(legend.title=element_blank())
emt_sig_umap
```

```{r}
ggsave(emt_sig_umap, filename="../figs/pan_cancer_sig_umap.png",
       width=8, height=7, dpi=300)
```

## RidgePlot
```{r}
library(ggridges)
```

```{r}
emt_sig_ridge <- ggplot(seurat@meta.data, aes(x=EMT_Sig1, y=SampleID, fill=stat(x))) +
  geom_density_ridges_gradient(alpha=0.5, quantile_lines=T, quantiles=2,
                               scale=2.5) +
  scale_fill_viridis_c(name="EMT Sig", option="A") +
  scale_x_continuous(limits=c(-0.1, 1.25)) +
  scale_y_discrete(expand = expand_scale(mult = c(0.01, 0.025))) +
  ylab("") + xlab("EMP Score") +
  theme_ridges(center_axis_labels = T) +
  theme(legend.position="none",
        axis.text.y=element_text(size=5, color="black"))
emt_sig_ridge
```

```{r}
ggsave(emt_sig_ridge, filename="../figs/pan_cancer_sig_ridge.pdf",
       width=5, height=13)
```

## Technical factors
```{r}
count_rna <- ggplot(dat, aes(UMAP1, UMAP2)) +
  geom_point(shape=16, size=0.2, aes(colour=log10(nCount_RNA)), alpha=0.5) +
  scale_colour_gradientn(colours=c(viridis::inferno(100))) +
  #scale_color_gradientn(colours=c('lightgrey', 'red')) +
  theme_void() +
  theme(legend.title=element_blank())
ggsave(count_rna, filename="../figs/pan_cancer_nCountRNA_umap.png",
       width=4.5, height=3.5, dpi=600)
count_rna
```

```{r}
feature_rna <- ggplot(dat, aes(UMAP1, UMAP2)) +
  geom_point(shape=16, size=0.2, aes(colour=nFeature_RNA), alpha=0.5) +
  scale_colour_gradientn(colours=c(viridis::inferno(100))) +
  #scale_color_gradientn(colours=c('lightgrey', 'red')) +
  theme_void() +
  theme(legend.title=element_blank())
ggsave(feature_rna, filename="../figs/pan_cancer_nFeatureRNA_umap.png",
       width=4.5, height=3.5, dpi=600)
feature_rna
```
```{r}
percent_mito <- ggplot(dat, aes(UMAP1, UMAP2)) +
  geom_point(shape=16, size=0.2, aes(colour=percent.mito), alpha=0.5) +
  scale_colour_gradientn(colours=c(viridis::inferno(100))) +
  #scale_color_gradientn(colours=c('lightgrey', 'red')) +
  theme_void() +
  theme(legend.title=element_blank())
ggsave(percent_mito, filename="../figs/pan_cancer_percentmito_umap.png",
       width=4.5, height=3.5, dpi=600)
percent_mito
```


# EMP correlation with technical factors
```{r}
seurat <- readRDS("../data/pan_cancer_epi.rds")
```

```{r}
tech_factors <- c("nCount_RNA", "percent.mito")
```


```{r}
getCor <- function(df){
  total_arch <- unique(paste0("A", unique(df$Archetype_Membership)))
  total_arch <- total_arch[order(total_arch)]
  cor_mat <- cor(df[,total_arch], df[,tech_factors])
}

cor_list <- lapply(meta_list, getCor)
```

```{r}
for(i in 1:length(cor_list)){
  rownames(cor_list[[i]]) <- paste0(names(meta_list)[i], "_", rownames(cor_list[[i]]))
}
```

```{r}
cor_list_merge <- do.call("rbind", cor_list)
```

```{r}
emp_programs <- readLines("../output/emp_program_list_broad.txt")
```

```{r}
cor_df <- as.data.frame(cor_list_merge)
cor_df$EMP <- rownames(cor_list_merge) %in% emp_programs
```

```{r}
mito_density <- ggplot(cor_df, aes(x=percent.mito)) +
  geom_density(aes(fill=EMP), color="black", alpha=0.75) +
  theme_classic()

count_density <- ggplot(cor_df, aes(x=nCount_RNA)) +
  geom_density(aes(fill=EMP), color="black", alpha=0.75) +
  theme_classic()

mito_density
count_density
```

Hm, not the greatest here. 

```{r}
emp_df <- filter(cor_df, EMP == T) %>%
  pivot_longer(col=-EMP, names_to = "Variable", values_to="Correlation")
```

```{r}
variable_plot <- ggplot(emp_df, aes(x=Variable, y=Correlation)) +
  geom_boxplot(color="black", fill="firebrick", alpha=0.75) +
  theme_classic() +
  theme(axis.text.x=element_text(size=12, angle=45, hjust=1, color="black"),
        axis.text.y=element_text(size=10, color="black"),
        axis.title=element_text(size=14))
variable_plot
```


```{r}
emp_df %>% group_by(Variable) %>% summarise(Median = median(Correlation))
```

# Coefficient ~ Cancer
Goal here to see if genes are associated with specific cancer types

```{r}
emt_genes <- readLines("../output/conserved_emt_signature.txt")
emt_programs <- readLines("../output/emp_program_list_broad.txt")
```


```{r}
cancer_dat <- program_scores %>% filter(Program %in% emt_programs)
cancer_dat$Cancer <- stringr::word(cancer_dat$Program, start=1, end=1, sep="_")
```

Hm, now, the challenge is that genes are not actually tested in all tumours. Do we only model genes in samples that they were explicitly tested or assume coef=0 if it's not detected. Actually, the problem of setting coef=0 is that it could artificially shrink variance. I think it's best to only go with the coefs we have.

```{r}
testCancer <- function(gene){
  df <- filter(cancer_dat, Gene == gene)
  res <- lm(df$coef ~ df$Cancer)
  res <- anova(res)
  res <- data.frame(Gene = gene,
                    pval = res$`Pr(>F)`[1])
}

cancer_res <- lapply(emt_genes, testCancer)
cancer_res <- do.call("rbind", cancer_res)
cancer_res$padj <- p.adjust(cancer_res$pval, method="BH")
```

```{r}
cancer_spec_genes <- cancer_res %>% filter(padj <= 0.05) %>% pull(Gene)
cancer_spec_genes
```

Now, we'll make a heatmap of genes x cancer type

```{r}
cancer_specific <- cancer_dat %>%
  filter(Gene %in% cancer_spec_genes) %>%
  group_by(Gene, Cancer) %>%
  summarise(Avg_coef = mean(coef)) %>%
  pivot_wider(names_from = Cancer,
              values_from = Avg_coef)

coef_mat <- as.matrix(cancer_specific[,2:ncol(cancer_specific)])
rownames(coef_mat) <- cancer_specific$Gene

```

```{r}
pheatmap::pheatmap(coef_mat,
                   clustering_method = "ward.D2",
                   color = colorRampPalette(rev(RColorBrewer::brewer.pal(7, "RdBu")))(100),
                   breaks = seq(-1, 1, length.out=101),
                   border_color="black",
                   filename = "~/Downloads/coef.png",
                   width=4, height=10)
```

