---
title: "Figure 5 Panels"
output: html_notebook
---

# Dependencies
```{r}
library(Seurat)
library(tidyverse)
library(factoextra)
library(progeny)
library(ggridges)
library(cowplot)
library(ggforce)
```

# Load data
```{r}
seurat <- readRDS("~/Data/single_cell_datasets/mix_seq/output/seurat_processed.rds")
meta_list <- readRDS(file="../../../output/master_metadata_mixseq.rds")
program_scores <- read.csv("../../../output/archetype_program_scores_mixseq.csv")
drug_sensitivity <- read.csv("../../../output/mixseq_emp_drug_sensitivity.csv", row.names=1)
summary_mat <- drug_sensitivity[,c("Drug", "Program", "Estimate")]
```

```{r}
summary_mat <- pivot_wider(summary_mat, names_from = Program, values_from = Estimate)
```

```{r}
drug_names <- summary_mat$Drug
summary_mat <- as.matrix(summary_mat[,2:ncol(summary_mat)])
rownames(summary_mat) <- drug_names
```

```{r}
seurat$CellLine_Full <- seurat$singlet_ID
```

```{r}
seurat@meta.data <- separate(seurat@meta.data, singlet_ID, into = c("CellLine", "Cancer"),
                             extra="merge")
```

# Fig 5A - PROGENy PCA w/ Trametinib effect sizes
## Test pathways
```{r}
#pathways <- colnames(progeny_mat)
sample_list <- names(meta_list)
pathways <- c("Androgen", "EGFR", "Estrogen", "Hypoxia", "JAK-STAT", "MAPK", "NFkB", "p53", "PI3K", "TGFb", "TNFa", "Trail", "VEGF", "WNT")

testPathway <- function(pathway, program, df){
  score_norm <- df[,program] / max(df[,program])
  test_summary <- summary(lm(df[,pathway] ~ score_norm))
  test_res <- data.frame(Coef = test_summary$coefficients[2,1],
                         pval = test_summary$coefficients[2,4],
                         Pathway = pathway)
  return(test_res)
}

runPathwayTests <- function(sampleID){
  print(paste0("Testing: ", sampleID))
  df <- meta_list[[sampleID]]
  total_arch <- unique(paste0("A", unique(df$Archetype_Membership)))
  total_arch <- total_arch[order(total_arch)]
  
  pathway_test_res <- list()
  for(i in 1:length(total_arch)){
    test_summary <- lapply(pathways, testPathway, program = total_arch[i], df = df)
    test_summary <- do.call("rbind", test_summary)
    test_summary$padj <- p.adjust(test_summary$pval, method="BH")
    test_summary$Program <- paste0(sampleID, "_", total_arch[i])
    pathway_test_res[[i]] <- test_summary
  }
  
  pathway_test_res <- do.call("rbind", pathway_test_res)
}
```

```{r}
pathway_scores <- lapply(sample_list, runPathwayTests)
pathway_scores <- do.call("rbind", pathway_scores)
```

## PCA of coefficients
```{r}
pathway_coef <- pathway_scores
pathway_coef$padj <- NULL
pathway_coef$pval <- NULL
pathway_coef <- pivot_wider(pathway_coef, names_from=Pathway, values_from=Coef)
pathway_coef <- as.data.frame(pathway_coef)
rownames(pathway_coef) <- pathway_coef$Program
pathway_coef$Program <- NULL
pathway_coef <- as.matrix(pathway_coef)
rownames(pathway_coef) <- make.names(rownames(pathway_coef))
```

```{r}
pca_res <- prcomp(pathway_coef, scale=F)
```

## Trametinib sensitivity

```{r}
tram_sensitivity <- drug_sensitivity %>%
  filter(Drug == "Trametinib")
tram_effect <- tram_sensitivity[match(rownames(pathway_coef), tram_sensitivity$Program), "Estimate"]

tram_effect[tram_effect < -0.15] <- -0.15
tram_effect[tram_effect > 0.15] <- 0.15
```

Re-order to make sure values get put on top
```{r}
tram_order <- rev(order(tram_effect))
pca_res$x <- pca_res$x[tram_order,]
```


```{r}
pca1 <- fviz_pca_biplot(pca_res, geom.ind="point", axes=c(1,2),
                col.ind = tram_effect[tram_order], repel=T, col.var = "black",
                ggtheme = theme_classic(),
                pointshape = 16, pointsize=2, title="MIX-seq: EMP-associated signalling")  +
  scale_color_gradientn(colours=colorRampPalette(rev(brewer.pal(7, "RdBu")))(100),
                        name="Trametinib effect\non EMP",
                        guide = guide_colorbar(ticks.colour = "black",
                                              frame.colour = "black",
                                              barwidth=0.75)) +
  theme(axis.text=element_text(size=10, colour='black'),
        axis.title=element_text(size=12),
        legend.text=element_text(size=12))

ggsave(pca1, filename="trametinib_sensitivity_pca.pdf", device=cairo_pdf(),
       width=4.5, height=3)

pca1
```


# Fig 5B - Boxplot of EMP

Look at trametinib effects. The Drug sensitivity data frame is only on EMP programs, so no need to subset it for EMP programs
```{r}
drug_sensitivity %>% 
  filter(Drug == "Trametinib") %>% 
  arrange(desc(Estimate))
```
Get genes associated with programs we choose from above 
```{r}
getGenes <- function(program){
  hits <- program_scores %>% filter(Program == program &
                                padj <= 0.05 &
                                coef >= 0.25) %>%
    pull(Gene)
}

loximvi_genes <- getGenes("LOXIMVI_A5")
msto211h_genes <- getGenes("MSTO211H_A6")
malme3m_genes <- getGenes("MALME3M_A6")
hec1b_genes <- getGenes("HEC1B_A3")
```

Set up scoring function
```{r}
scoreCells <- function(cell_line, gene_set){
  seurat_subset <- subset(seurat, 
                          subset = CellLine == cell_line & 
                            Drug %in% c("Trametinib", "Untreated", "DMSO"))
  seurat_subset <- AddModuleScore_UCell(seurat_subset, features = list(EMP_Sig=gene_set),
                                  name = "")
  return(seurat_subset@meta.data)
}
```

```{r}
loximvi <- scoreCells("LOXIMVI", loximvi_genes)
loximvi$Drug <- factor(loximvi$Drug, levels=c("DMSO", "Trametinib"))

msto211h <- scoreCells("MSTO211H", msto211h_genes)
msto211h$Drug <- factor(msto211h$Drug, levels=c("DMSO", "Trametinib"))

malme3m <- scoreCells("MALME3M", malme3m_genes)
malme3m$Drug <- factor(malme3m$Drug, levels=c("DMSO", "Trametinib"))

hec1b <- scoreCells("HEC1B", hec1b_genes)
hec1b$Drug <- factor(hec1b$Drug, levels=c("DMSO", "Trametinib"))
```

Plot
```{r}
plotSample <- function(dat, colour){
  dat_plot <- ggplot(dat, aes(Drug, EMP_Sig)) +
    geom_sina(aes(color=Drug), alpha=0.75, shape=16, size=1) +
    stat_summary(fun.y="mean",geom="crossbar", width=0.5) +
    scale_color_manual(values = c("grey", colour)) +
    ylab("EMP UCell") + xlab("") +
    ggtitle(paste0(unique(dat$CellLine),
                   " - ",
                   Hmisc::capitalize(tolower(unique(dat$Cancer))))) +
    theme_classic() +
    theme(legend.position="none",
          axis.text.x=element_text(size=12, angle=45, hjust=1, color="black"),
          axis.text.y=element_text(size=10, color="black"),
          axis.title = element_text(size=12),
          plot.title = element_text(size=10, color="black"))
  dat_plot
}

loximvi_plot <- plotSample(loximvi, RColorBrewer::brewer.pal(7, "RdBu")[1])
msto211h_plot <- plotSample(msto211h, RColorBrewer::brewer.pal(7, "RdBu")[1])
malme3m_plot <- plotSample(malme3m, RColorBrewer::brewer.pal(7, "RdBu")[7])
hec1b_plot <- plotSample(hec1b, RColorBrewer::brewer.pal(7, "RdBu")[7])


drug_plots <- cowplot::plot_grid(loximvi_plot, msto211h_plot, malme3m_plot, hec1b_plot, ncol=2,
                                 align = "hv")

cowplot::save_plot(drug_plots, filename="trametinib_sample_lines.pdf",
                   base_width = 4.2, base_height=4.0)
```

# Fig 5C - Cook lines pathway ~ EMP
```{r}
a549 <- readRDS("~/Projects/emt_dynamics/data/A549_Kinase_Untreated.rds")
du145 <- readRDS("~/Projects/emt_dynamics/data/DU145_Kinase_Untreated.rds")
mcf7 <- readRDS("~/Projects/emt_dynamics/data/MCF7_Kinase_Untreated.rds")
ovca420 <- readRDS("~/Projects/emt_dynamics/data/OVCA420_Kinase_Untreated.rds")

seurat <- merge(a549, list(du145, mcf7, ovca420))

meta_list <- readRDS("../../../output/master_metadata_cook_kinase.rds")
emt_programs <- readLines("../../../output/cook_emt_programs.txt")
program_scores <- read.csv("../../../output/archetype_program_scores_cook_kinase.csv")
drug_sensitivity <- read.csv("../../../output/cook_kinase_emp_drug_sensitivity.csv")
```

Because we have >1 EMT program per line, I'll just do one per line, choosing the one that correlates best with EMT programs

```{r}
emt_programs <- c("A549_A3", "OVCA420_A4", "DU145_A5", "MCF7_A2")
```

```{r}
plotPathway <- function(cell_line, program){
  dat <- meta_list[[cell_line]]
  dat[,program]<- dat[,program] / max(dat[,program])
  dat_plot <- ggplot(dat, aes_string(x=program, y="TGFb")) +
    geom_point(size=1, shape=16, alpha=0.75, color="black") +
    geom_smooth(color="firebrick") +
    xlab("EMP Program") + ylab("TGFB Activity") +
    ggtitle(cell_line) +
    theme_classic() +
    theme(axis.text=element_text(size=10, colour='black'),
          axis.title=element_text(size=12),
          plot.title = element_text(size=12))
  dat_plot
}

a549_plot <- plotPathway("A549", "A3")
du145_plot <- plotPathway("DU145", "A5")
mcf7_plot <- plotPathway("MCF7", "A2")
ovca420_plot <- plotPathway("OVCA420", "A4")
```



# Fig 5D - TGFBR1 inhibition on each program
```{r}
getGenes <- function(program){
  hits <- program_scores %>% filter(Program == program &
                                padj <= 0.05 &
                                coef >= 0.25) %>%
    pull(Gene)
}

a549_genes <- getGenes("A549_A3")
du145_genes <- getGenes("DU145_A5")
mcf7_genes <- getGenes("MCF7_A2")
ovca420_genes <- getGenes("OVCA420_A4")
```

Set function to score cells by sample-specific EMP
```{r}
scoreCells <- function(cell_line, gene_set){
  seurat_subset <- subset(seurat, 
                          subset = CellLine == cell_line & 
                            Drug %in% c("Uninhibited_Untreated", "LY364947"))
  seurat_subset <- AddModuleScore_UCell(seurat_subset, features = list(EMP_Sig=gene_set),
                                  name = "")
  return(seurat_subset@meta.data)
}

a549_scores <- scoreCells("A549", a549_genes)
du145_scores <- scoreCells("DU145", du145_genes)
mcf7_scores <- scoreCells("MCF7", mcf7_genes)
ovca420_scores <- scoreCells("OVCA420", ovca420_genes)
```

```{r}
plotSample <- function(dat){
  dat$Drug[dat$Drug == "Uninhibited_Untreated"] <- "Untreated"
  dat$Drug <- factor(dat$Drug, levels=c("Untreated", "LY364947"))
  dat_plot <- ggplot(dat, aes(Drug, EMP_Sig)) +
    geom_sina(aes(color=Drug), alpha=0.75, shape=16, size=1) +
    stat_summary(fun.y="mean",geom="crossbar", width=0.5) +
    scale_color_manual(values = c("grey", RColorBrewer::brewer.pal(7, "RdBu")[7])) +
    ylab("EMP Program") + xlab("") +
    ggtitle("TGFBR1 Inhibition") +
    theme_classic() +
    theme(legend.position="none",
          axis.text.x=element_text(size=12, angle=45, hjust=1, color="black"),
          axis.text.y=element_text(size=12, color="black"),
          axis.title = element_text(size=12),
          plot.title = element_text(size=12, color="black"))
  dat_plot
}

a549_inhibition <- plotSample(a549_scores)
du145_inhibition <- plotSample(du145_scores)
mcf7_inhibition <- plotSample(mcf7_scores)
ovca420_inhibition <- plotSample(ovca420_scores)
```


```{r}
tgfb_plots <- plot_grid(a549_plot, du145_plot, mcf7_plot, ovca420_plot,
                        a549_inhibition, du145_inhibition, mcf7_inhibition, ovca420_inhibition,
                        ncol=4,
                        align="v")
cowplot::save_plot(tgfb_plots, filename="tgfb_inhibition_plots.pdf",
                   base_width=9, base_height=5)

tgfb_plots
```


## Some statistics
### Pathway activity
```{r}
#pathways <- colnames(progeny_mat)
pathways <- c("Androgen", "EGFR", "Estrogen", "Hypoxia", "JAK-STAT", "MAPK", "NFkB", "p53", "PI3K", "TGFb", "TNFa", "Trail", "VEGF", "WNT")

testPathway <- function(pathway, program, df){
  score_norm <- df[,program] / max(df[,program])
  test_summary <- summary(lm(df[,pathway] ~ score_norm))
  test_res <- data.frame(Coef = test_summary$coefficients[2,1],
                         pval = test_summary$coefficients[2,4],
                         Pathway = pathway)
  return(test_res)
}

runPathwayTests <- function(sampleID){
  print(paste0("Testing: ", sampleID))
  df <- meta_list[[sampleID]]
  total_arch <- unique(paste0("A", unique(df$Archetype_Membership)))
  total_arch <- total_arch[order(total_arch)]
  
  pathway_test_res <- list()
  for(i in 1:length(total_arch)){
    test_summary <- lapply(pathways, testPathway, program = total_arch[i], df = df)
    test_summary <- do.call("rbind", test_summary)
    test_summary$padj <- p.adjust(test_summary$pval, method="BH")
    test_summary$Program <- paste0(sampleID, "_", total_arch[i])
    pathway_test_res[[i]] <- test_summary
  }
  
  pathway_test_res <- do.call("rbind", pathway_test_res)
}
```

```{r}
pathway_scores <- lapply(unique(seurat$CellLine), runPathwayTests)
pathway_scores <- do.call("rbind", pathway_scores)
pathway_scores %>% filter(Program %in% emt_programs &
                            Pathway == "TGFb")
```

### Drug effect
```{r}
drug_sensitivity %>% filter(Program %in% emt_programs & Drug == "LY364947")
```












