---
title: "Figure S1 - Processing pipeline"
output: html_notebook
---

```{r}
library(Seurat)
library(cowplot)
library(tidyverse)
```

```{r}
seurat <- readRDS("~/Projects/emt_programs/output/tumour_annotations/colorectal_lee_seurat.rds")
```

```{r}
p1 <- DimPlot(seurat, group.by="seurat_clusters", label=T, shuffle=T, pt.size=0.1) + NoAxes() + NoLegend()
p2 <- DimPlot(seurat, group.by="Sample", shuffle=T,pt.size=0.1) + NoAxes() + NoLegend()
p3 <- DimPlot(seurat, group.by="CellType", label=T, shuffle=T, pt.size=0.1) + NoAxes() + NoLegend()
```

```{r}
cowplot::plot_grid(p1, p2, p3, ncol=3)
```

# Function
```{r}
plotData <- function(seurat, title){
  p1 <- DimPlot(seurat, group.by="seurat_clusters", label=T, 
                shuffle=T, pt.size=0.1, raster=T) + 
    NoAxes() + NoLegend() + ggtitle("Clusters") +
    theme(plot.title = element_text(color="black", size=14, face="plain"))
  p2 <- DimPlot(seurat, group.by="Sample", 
                shuffle=T,pt.size=0.1, raster=T) + 
    NoAxes() + NoLegend()  + ggtitle("Sample") +
    theme(plot.title = element_text(color="black", size=14, face="plain"))
  p3 <- DimPlot(seurat, group.by="CellType", label=T, 
                shuffle=T, pt.size=0.1, raster=T) + 
    NoAxes() + NoLegend() + ggtitle("Cell type (singleR)") +
    theme(plot.title = element_text(color="black", size=14, face="plain"))
  
  plot_row <- cowplot::plot_grid(p1,p2,p3, ncol=3)
  
  #Make title
  title_plot <- ggdraw() +
    draw_label(title,
               x=0,
               hjust=0) +
    theme(plot.margin = margin(0,0,0,7))
  
  #Plot all
  dat_plot <- cowplot::plot_grid(title_plot, plot_row, ncol=1, rel_heights = c(0.1, 1))
  
 save_plot(filename = paste0("~/Projects/emt_programs/fig_panels/revised_figures/figS1/", 
                             title, ".pdf"),
            dat_plot,
            base_width = 12, base_height=4, device="pdf")
}
```

# Colorectal Lee
```{r}
seurat <- readRDS("../../../output/tumour_annotations/colorectal_lee_seurat.rds")
plotData(seurat, "Colorectal - Lee et al")
```

# Colorectal Lee - KUL
```{r}
seurat <- readRDS("../../../output/tumour_annotations/colorectal_lee_KUL_seurat.rds")
plotData(seurat, "Colorectal - KUL - Lee et al")
```

# Colorectal - Uhlitz et al 
```{r}
seurat <- readRDS("../../../output/tumour_annotations/colorectal_Uhlitz_seurat.rds")
seurat$Sample <- seurat$orig.ident
plotData(seurat, "Colorectal - Uhlitz et al")
```

# Colorectal - Qian et al

```{r}
seurat <- readRDS("../../../output/tumour_annotations/colorectal_Qian_seurat.rds")
seurat$Sample <- seurat$orig.ident
plotData(seurat, "Colorectal - Qian et al")
```

# Gastric - Sathe et al

```{r}
seurat <- readRDS("../../../output/tumour_annotations/gastric_sathe_seurat.rds")
seurat$Sample <- seurat$orig.ident
plotData(seurat, "Gastric - Sathe et al")
```

# Lung - Kim et al - GSE131907

```{r}
seurat <- readRDS("../../../output/tumour_annotations/lung_kim_seurat.rds")
plotData(seurat, "Lung - Kim et al")
```

# Lung - Lambrechts et al

```{r}
seurat <- readRDS("../../../output/tumour_annotations/lung_lambrechts_seurat.rds")
seurat$Sample <- seurat$orig.ident
plotData(seurat, "Lung - Lambrechts et al")
```

# Lung - Laughey et al - GSE123904

```{r}
seurat <- readRDS("../../../output/tumour_annotations/lung_laughney_seurat.rds")
seurat$Sample <- seurat$Patient
plotData(seurat, "Lung - Laughney et al")
```

# Lung - Qian et al - http://blueprint.lambrechtslab.org

```{r}
seurat <- readRDS("../../../output/tumour_annotations/lung_qian_seurat.rds")
seurat$Sample <- seurat$Patient
plotData(seurat, "Lung - Qian et al")
```

# Lung - Wu et al

```{r}
seurat <- readRDS("../../../output/tumour_annotations/lung_wu_seurat.rds")
plotData(seurat, "Lung - Wu et al")
```

# Ovarian - Geistlinger et al - GSE154600

```{r}
seurat <- readRDS("../../../output/tumour_annotations/ovarian_geistlinger_seurat.rds")
seurat$Sample <- seurat$orig.ident
plotData(seurat, "Ovarian - Geistlinger et al")
```

# Ovarian - Qian et al - http://blueprint.lambrechtslab.org

```{r}
seurat <- readRDS("../../../output/tumour_annotations/ovarian_qian_seurat.rds")
seurat$Sample <- seurat$Patient
plotData(seurat, "Ovarian - Qian et al")
```

# PDAC - Steele et al - GSE155698

```{r}
seurat <- readRDS("../../../output/tumour_annotations/pdac_steele_seurat.rds")
plotData(seurat, "PDAC - Steele et al")
```

# Breast - Wu et al - ENA PRJEB35405

```{r}
seurat <- readRDS("../../../output/tumour_annotations/breast_wu_seurat.rds")
seurat$Sample <- seurat$Patient
plotData(seurat, "Breast - Wu et al")
```

# Breast - Qian et al - http://blueprint.lambrechtslab.org

```{r}
seurat <- readRDS("../../../output/tumour_annotations/breast_qian_seurat.rds")
seurat$Sample <- seurat$Patient
plotData(seurat, "Breast - Qian et al")
```

# Breast - Bassez et al - http://biokey.lambrechtslab.org/

## Cohort 1

```{r}
seurat <- readRDS("../../../output/tumour_annotations/breast_bassez_cohort1_seurat.rds")
seurat$Sample <- seurat$Patient
plotData(seurat, "Breast - Cohort1 - Bassez et al")
```

## Cohort 2

```{r}
seurat <- readRDS("../../../output/tumour_annotations/breast_bassez_cohort2_seurat.rds")
seurat$Sample <- seurat$Patient
plotData(seurat, "Breast - Cohort2 - Bassez et al")
```

# SCC - Ji et al - GSE144236

```{r}
seurat <- readRDS("../../../output/tumour_annotations/scc_ji_seurat.rds")
seurat$Sample <- seurat$Patient
plotData(seurat, "SCC - Ji et al")
```

# Nasopharyngeal Carcinoma - Chen et al - GSE150430

```{r}
seurat <- readRDS("../../../output/tumour_annotations/npc_chen_seurat.rds")
seurat$Sample <- seurat$Patient
plotData(seurat, "NPC - Chen et al")
```

