---
title: "Figure 4 Panels"
output: html_notebook
---

# Dependencies
```{r}
library(Seurat)
library(tidyverse)
library(factoextra)
library(progeny)
library(ggrepel)
```

```{r}
meta_list <- readRDS("../../../output/master_metadata.rds")
sample_list <- names(meta_list)
emp_programs <- readLines("../../../output/emp_program_list_broad.txt")
emt_sig <- readLines("../../../output/conserved_emt_signature.txt")
```



# Fig 4B - Ligand count
Data from nichenet.Rmd

```{r}
niche_res <- readRDS("../../../output/nichenet_res.rds")
```

Get top 10 ligands for each sample, count how often each ligand comes up
```{r}
good_ligands <- niche_res %>%
  select(c("Sample", "pearson", "Gene")) %>%
  unique() %>%
  group_by(Sample) %>%
  top_n(10, pearson) %>%
  ungroup() %>%
  group_by(Gene) %>%
  summarise(Count = length(Sample)) %>%
  arrange(desc(Count)) %>%
  mutate(Rank = 1:nrow(.))
good_ligands
```


```{r}
normal12 <- c("#ebac23", #EMP
              "#008cf9",#Cancer
              "#b80058", #Fib
              "#B24502", #Smooth muscle
              "#00bbad", #Endo
              "#d163e6", #Macro
              "#FF9287", #Mono
              "#006e00", #B
              "#5954D6", #T
              "#00C6F8", #NK
              "#878500", #DC
              "#00A76C", #Platelets
              "#bdbdbd") #Unkown
```


```{r}
good_ligands$Label <- ""
good_ligands$Label[1:25] <- good_ligands$Gene[1:25]


## Add color by highest expressing cell type
niche_summary <- niche_res %>%
  filter(Gene %in% unique(good_ligands$Gene)) %>%
  group_by(CellType, Gene) %>%
  summarise(Exp = mean(AvgExp, na.rm = T),
            Pct = mean(pct, na.rm=T)) %>%
  ungroup() %>%
  group_by(Gene) %>%
  filter(Exp == max(Exp))

merge_dat <- left_join(good_ligands, niche_summary, by="Gene")
merge_dat$CellType <- factor(merge_dat$CellType, 
                             levels = c("Cancer cells_EMP", "Cancer cells", "Fibroblasts",
                                        "Smooth_muscle_cells", "Endothelial_cells", 
                                        "Macrophage", "Monocyte", "B_cell", 
                                        "T_cells", "NK_cell", "DC", "Platelets", "Unknown"))


ligand_freq_plot <- ggplot(merge_dat, aes(Rank, Count, label=Label)) +
  geom_point(size=1, color="black") +
  geom_text_repel(aes(color=CellType),
                  max.overlaps = 1000,
                  min.segment.length=0,
                  direction="y",
                  nudge_x=125,
                  size=2.8,
                  hjust=0,
                  segment.size=0.2,
                  segment.alpha=0.75,
                  segment.color="black") +
  xlab("Ligands ranked by frequency") + ylab("Number of tumours with\npredicted signalling") + 
  scale_y_continuous(breaks=c(0,20, 40, 60, 80, 100), limits=c(0,100)) +
  scale_color_manual(values=normal12,
                     name="Highest expressing\ncell type") +
  theme_classic() +
  theme(axis.text = element_text(size=10, color="black"),
        axis.title = element_text(size=12),
        legend.key.size = unit(0.15, "cm"),
        legend.text=element_text(size=10),
        legend.title=element_text(size=12))

ggsave(ligand_freq_plot, filename="nichenet_top_ligand_freq.pdf",
       width=6, height=3, device=cairo_pdf())
```


# Fig 4C - Ligand dot plot
```{r}
ligands <- good_ligands$Gene[1:100]
```


```{r}
niche_summary <- niche_res %>%
  filter(Gene %in% ligands) %>%
  group_by(CellType, Gene) %>%
  summarise(Exp = mean(AvgExp, na.rm = T),
            Pct = mean(pct, na.rm=T))
```


I'm going to make a dotplot, but I'd like the ligands/cell types to be ordered in a visually appealing way. I'll make a clustered heatmap, steal the order, and then use it on this
```{r}
niche_tmp <- niche_summary %>%
  select(c("CellType", "Gene", "Exp")) %>%
  pivot_wider(names_from="Gene", values_from="Exp")
niche_mat <- as.matrix(niche_tmp[,2:ncol(niche_tmp)])
rownames(niche_mat) <- niche_tmp$CellType

niche_heat <- pheatmap::pheatmap(niche_mat,
                                 silent=T,
                                 clustering_method = "ward.D2")
```


```{r}
#Aesthetic things
niche_dot <- niche_summary
niche_dot$Exp[niche_dot$Exp > 1] <- 1
niche_dot$Pct[niche_dot$Pct > 0.75] <- 0.75
niche_dot$Pct <- niche_dot$Pct * 100
niche_dot$CellType <- factor(niche_dot$CellType,
                             levels = rev(c("Cancer cells_EMP", "Cancer cells", "Fibroblasts",
                                        "Smooth_muscle_cells", "Endothelial_cells", 
                                        "Macrophage", "Monocyte", "B_cell", 
                                        "T_cells", "NK_cell", "DC", "Platelets", "Unknown")))
niche_dot$Gene <- factor(niche_dot$Gene,
                             levels = colnames(niche_mat)[niche_heat$tree_col$order])

#Plot
dot <- ggplot(niche_dot, aes(x=Gene, y=CellType)) +
  geom_point(aes(size = Pct, fill = Exp), alpha=0.8, color="black", shape=21) +
  scale_size("% Exp", range = c(0,5)) +
  scale_fill_gradientn(colours = viridisLite::inferno(100),
                       limits = c(0,1),
                       guide = guide_colorbar(ticks.colour = "black",
                                              frame.colour = "black"),
                       name = "Avg Exp\nlog(TP10k)") +
  ylab("") + xlab("") +
  theme_bw() +
  theme(axis.text.x = element_text(size=9, angle=45, hjust=1, color="black"),
        axis.text.y = element_text(size=12, color="black"))

#Save
ggsave(dot, filename="nichnet_dotplot.pdf",
       device = cairo_pdf(),
       width=18, height=4.25)
```

# Fig 4D - PCA of Progeny coeffients
See emp_regulation.Rmd for the script that ran progeny on all the samples. Took a while and saved data into meta_list

First, test each program for changes in pathways
```{r}
#pathways <- colnames(progeny_mat)
pathways <- c("Androgen", "EGFR", "Estrogen", "Hypoxia", "JAK-STAT", "MAPK", "NFkB", "p53", "PI3K", "TGFb", "TNFa", "Trail", "VEGF", "WNT")

testPathway <- function(pathway, program, df){
  score_norm <- df[,program] / max(df[,program])
  test_summary <- summary(lm(df[,pathway] ~ score_norm))
  test_res <- data.frame(Coef = test_summary$coefficients[2,1],
                         pval = test_summary$coefficients[2,4],
                         Pathway = pathway)
  return(test_res)
}

runPathwayTests <- function(sampleID){
  print(paste0("Testing: ", sampleID))
  df <- meta_list[[sampleID]]
  total_arch <- unique(paste0("A", unique(df$Archetype_Membership)))
  total_arch <- total_arch[order(total_arch)]
  
  pathway_test_res <- list()
  for(i in 1:length(total_arch)){
    test_summary <- lapply(pathways, testPathway, program = total_arch[i], df = df)
    test_summary <- do.call("rbind", test_summary)
    test_summary$padj <- p.adjust(test_summary$pval, method="BH")
    test_summary$Program <- paste0(sampleID, "_", total_arch[i])
    pathway_test_res[[i]] <- test_summary
  }
  
  pathway_test_res <- do.call("rbind", pathway_test_res)
}
```

```{r}
pathway_scores <- lapply(sample_list, runPathwayTests)
pathway_scores <- do.call("rbind", pathway_scores)
```

Build coefficient matrix

```{r}
pathway_coef <- pathway_scores
pathway_coef$padj <- NULL
pathway_coef$pval <- NULL
pathway_coef <- pivot_wider(pathway_coef, names_from=Pathway, values_from=Coef)
pathway_coef <- as.data.frame(pathway_coef)
rownames(pathway_coef) <- pathway_coef$Program
pathway_coef$Program <- NULL
pathway_coef <- as.matrix(pathway_coef)
```

## KUL01 example
### Arch program PCA
```{r}
seurat <- readRDS("../../../data/pan_cancer_epi.rds")
```

```{r}
sampleID <- "Colorectal_Lee_KUL01"
cells <- rownames(meta_list[[sampleID]])
seurat_subset <- subset(seurat, cells = cells)
seurat_subset <- SCTransform(seurat_subset, vars.to.regress="percent.mito")
seurat_subset <- RunPCA(seurat_subset, verbose=F)

dat <- meta_list[[sampleID]]
dat$PC1 <- Embeddings(seurat_subset, "pca")[,1]
dat$PC2 <- Embeddings(seurat_subset, "pca")[,2]
dat$A1 <- dat$A1 / max(dat$A1)
dat$A2 <- dat$A2 / max(dat$A2)
dat$A3 <- dat$A3 / max(dat$A3)
dat$A4 <- dat$A4 / max(dat$A4)
```

```{r}
df <- dat %>%
  select(PC1, PC2, A1, A2, A3, A4) %>%
  pivot_longer(cols = -c(PC1, PC2),
               names_to = "Program",
               values_to = "Score")
```

```{r}
pca_programs <- ggplot(df, aes(PC1, PC2)) +
  geom_point(size=0.5, shape=16, aes(color=Score)) +
  scale_color_gradientn(colors=viridis::mako(100), #colours = c("lightgrey", "red"),
                        name="Program\nscore",
                        guide = guide_colorbar(ticks.colour = "black",
                                              frame.colour = "black",
                                              barwidth=0.75)) +
  facet_wrap(~Program, ncol =4) +
  theme_bw() +
  theme(axis.text=element_blank(),#element_text(size=10, color="black"),
        axis.title=element_blank(),#element_text(size=12),
        axis.ticks=element_blank(),
        panel.grid.major = element_blank(),
        panel.grid.minor = element_blank(),
        legend.title=element_text(size=12),
        legend.text=element_text(size=12),
        strip.background = element_blank(),
        strip.text = element_text(size=12, color="black"),
        panel.border = element_rect(size=0.5))
pca_programs

pca_programs <- ggplot(df, aes(PC1, PC2)) +
  geom_point(size=0.5, shape=16, aes(color=Score)) +
  scale_color_gradientn(colors=viridis::mako(100), #colours = c("lightgrey", "red"),
                        name="Program\nscore",
                        guide = guide_colorbar(ticks.colour = "black",
                                              frame.colour = "black",
                                              barwidth=0.75)) +
  facet_wrap(~Program, ncol =4) +
  theme_void() +
  theme(legend.title=element_text(size=12),
        legend.text=element_text(size=12),
        strip.background = element_blank(),
        strip.text = element_text(size=12, color="black"))
pca_programs
```

```{r}
ggsave(pca_programs, filename="kul01_program_pca.pdf",
       width=6.5, height=2, device=cairo_pdf())
```


### Signalling activity
```{r}
kul01_programs <- unique(grep("KUL01", pathway_scores$Program, value=T))
coef_kul01 <- pathway_coef[kul01_programs,]
rownames(coef_kul01) <- c("A1", "A2", "A3", "A4")
```

```{r}
pheatmap::pheatmap(coef_kul01,
                   border_color = "black",
                   color = colorRampPalette(rev(RColorBrewer::brewer.pal(7, "RdBu")))(100),
                   clustering_method="ward.D2",
                   cluster_rows=F,
                   treeheight_col = 0,
                   #angle_col = 45,
                   legend=F,
                   filename="kul01_progeny_heatmap_horizontal.png",
                   width=3, height=1.65, dpi=600)
```

### TNFa
```{r}
df <- meta_list[["Colorectal_Lee_KUL01"]]
df$A2 <- df$A2 / max(df$A2)
tnfa_plot <- ggplot(df, aes(x=A2, y=TNFa)) +
  geom_point(shape=16, size=0.5, color="firebrick") +
  geom_smooth(color="black") +
  ggtitle("KUL01 - TNFa activity") +
  xlab("A2 (EMP) score") + ylab("TNFa activity") +
  theme_classic() +
  theme(axis.text=element_text(size=10, color="black"),
          axis.title=element_text(size=12),
          plot.title = element_text(size=12))

ggsave(tnfa_plot, filename="kul01_tnf.pdf",
       width=2.75, height=2.15, device=cairo_pdf())
```


## PCA of signalling coefficients
```{r}
pca_res <- prcomp(pathway_coef, scale=F)
```

## Colored by cancer, but only EMP programs
Factoextra provides some nice plotting functionality w/ ggplot2. I think I'll just use its plots

```{r}
program_list <- data.frame(Program = rownames(pathway_coef),
                           EMP = ifelse(rownames(pathway_coef) %in% emp_programs, "EMP", "Other"),
                           Cancer = gsub("_.*$", "", rownames(pathway_coef))) #remove all characters after the first underscore
program_list <- arrange(program_list, desc(EMP))
program_list$Cancer[which(program_list$EMP == "Other")] <- "Z"
cols <- c("#ebac23", "#b80058", "#008cf9", "#006e00", "#00bbad", "#d163e6", "#b24502", "#ff9287", "lightgrey")

pca_res$x <- pca_res$x[program_list$Program,]
```


```{r}
pca1 <- fviz_pca_biplot(pca_res, geom.ind="point", axes=c(1,2),
                col.ind = program_list$Cancer, col.var = "black",
                repel=T, ggtheme = theme_classic(),
                pointshape = 19, pointsize=1.2, title="PCA of all model coefficients") + #oversize points -- remove stroke in Designer
  scale_color_manual(values=cols) +
  guides(colour = guide_legend(override.aes = list(size=4))) +
  theme(axis.text=element_text(size=12, colour='black'),
        axis.title=element_text(size=12),
        legend.text=element_text(size=12))

ggsave(pca1, filename="progeny_pca_cancer.pdf", device=cairo_pdf(),
       width=6, height=4)

pca1
```


# Fig 4D - MAPK / Hypoxia difference

Will make use of PCA results from Fig 4A above

## PCA component
```{r}
mapk_programs <- rownames(pca_res$x)[rownames(pca_res$x) %in% emp_programs &
                                       pca_res$x[,2] < -1 &
                                       pca_res$x[,1] > 0]
#mapk_programs <- substr(mapk_programs, 1, nchar(mapk_programs)-3) %>% unique()

stat_programs <- rownames(pca_res$x)[rownames(pca_res$x) %in% emp_programs &
                                       pca_res$x[,2] > 1 &
                                       pca_res$x[,1] > 0]
#stat_programs <- substr(stat_programs, 1, nchar(stat_programs)-3) %>% unique()
```

```{r}
program_list$Group <- "Z"
program_list$Group[program_list$Program %in% mapk_programs] <- "MAPK"
program_list$Group[program_list$Program %in% stat_programs] <- "STAT"

cols <- RColorBrewer::brewer.pal(3, "Dark2")
cols[3] <- "lightgrey"
```


```{r}
pca1 <- fviz_pca_biplot(pca_res, geom.ind="point", axes=c(1,2),
                col.ind = program_list$Group, col.var = "black",
                repel=T, ggtheme = theme_classic(),
                pointshape = 19, pointsize=1.2) + #oversize points -- remove stroke in Designer
  scale_color_manual(values=cols) +
  guides(colour = guide_legend(override.aes = list(size=4))) +
  theme(axis.text=element_text(size=10, colour='black'),
        axis.title=element_text(size=12),
        legend.position="none")

ggsave(pca1, filename="progeny_mapk_stat.pdf", device=cairo_pdf(),
       width=4.25, height=4)

pca1
```

## Compare ligands
```{r}
mapk_programs <- rownames(pca_res$x)[rownames(pca_res$x) %in% emp_programs &
                                       pca_res$x[,2] < -1 &
                                       pca_res$x[,1] > 0]
mapk_programs <- substr(mapk_programs, 1, nchar(mapk_programs)-3) %>% unique()

stat_programs <- rownames(pca_res$x)[rownames(pca_res$x) %in% emp_programs &
                                       pca_res$x[,2] > 1 &
                                       pca_res$x[,1] > 0]
stat_programs <- substr(stat_programs, 1, nchar(stat_programs)-3) %>% unique()
```

```{r}
mapk_ligands <- niche_res %>%
  filter(Sample %in% mapk_programs) %>%
  select(c("Sample", "pearson", "Gene")) %>%
  unique() %>%
  group_by(Sample) %>%
  top_n(20, pearson) %>% #results are actually a bit cleaner if you go with top 10
  ungroup()



mapk_ligand_count <- mapk_ligands %>% group_by(Gene) %>%
  summarise(Count = length(Sample)) %>%
  arrange(desc(Count)) %>%
  mutate(Rank = 1:nrow(.))
mapk_ligand_count$Count <- mapk_ligand_count$Count / length(mapk_programs)
#The above line accounts for differences in the number of MAPK and STAT-associated programs. Convert to proportion


stat_ligands <- niche_res %>%
  filter(Sample %in% stat_programs) %>%
  select(c("Sample", "pearson", "Gene")) %>%
  unique() %>%
  group_by(Sample) %>%
  top_n(20, pearson) %>% #results are actually a bit cleaner if you go with top 10
  ungroup()



stat_ligand_count <- stat_ligands %>% group_by(Gene) %>%
  summarise(Count = length(Sample)) %>%
  arrange(desc(Count)) %>%
  mutate(Rank = 1:nrow(.))
stat_ligand_count$Count <- stat_ligand_count$Count / length(stat_programs) 
```

```{r}
#common_ligands <- intersect(mapk_ligand_count, stat_ligand_count)



```

## Plot
Add colour by cell type with highest expression maybe??
```{r}
merge_dat <- left_join(mapk_ligand_count, stat_ligand_count, by="Gene")
merge_dat <- na.omit(merge_dat)
merge_dat$CountDiff <- merge_dat$Count.x - merge_dat$Count.y #The difference in proportion
merge_dat <- arrange(merge_dat, desc(CountDiff))
merge_dat$Rank <- 1:nrow(merge_dat)


mapk_ligands <- c(merge_dat %>% filter(CountDiff > 0.075) %>% pull(Gene)) 
stat_ligands <- c(merge_dat %>% filter(CountDiff < -0.075) %>% pull(Gene))

merge_dat$mapk_Label <- ""
merge_dat$mapk_Label[match(mapk_ligands, merge_dat$Gene)] <- mapk_ligands

merge_dat$stat_Label <- ""
merge_dat$stat_Label[match(stat_ligands, merge_dat$Gene)] <- stat_ligands

# Get expression info for ligands
niche_summary <- niche_res %>%
  filter(Gene %in% unique(merge_dat$Gene)) %>%
  group_by(CellType, Gene) %>%
  summarise(Exp = mean(AvgExp, na.rm = T),
            Pct = mean(pct, na.rm=T)) %>%
  ungroup() %>%
  group_by(Gene) %>%
  filter(Exp == max(Exp))

merge_dat <- left_join(merge_dat, niche_summary, by="Gene")
merge_dat$CellType <- factor(merge_dat$CellType, 
                             levels = c("Cancer cells_EMP", "Cancer cells", "Fibroblasts",
                                        "Smooth_muscle_cells", "Endothelial_cells", 
                                        "Macrophage", "Monocyte", "B_cell", 
                                        "T_cells", "NK_cell", "DC", "Platelets", "Unknown"))

# Plot
diff_plot <- ggplot(merge_dat, aes(x=Rank, y=CountDiff)) +
  geom_point(aes(fill=CellType, size=Exp), colour="black", shape=21, alpha=0.6) +
  geom_hline(yintercept = c(-0.075, 0.075), linetype=2) +
  geom_text_repel(aes(label = mapk_Label, color=CellType),
                  nudge_x = 50,
                  direction = "y",
                  force=0.5,
                  ylim = c(0.075,NA),
                  hjust= 0,
                  size=3,
                  segment.size = 0.2,
                  segment.color="black") +
  geom_text_repel(aes(label = stat_Label, color=CellType),
                  nudge_x = -70,
                  force=1,
                  direction = "y",
                  ylim = c(NA,-0.075),
                  hjust= 1,
                  size=2.75,
                  segment.size = 0.2,
                  segment.color="black") +
  scale_y_continuous(expand = c(0, 0), breaks=c(-0.3, -0.2, -0.1, 0, 0.1, 0.2, 0.3), limits=c(-0.3, 0.3)) +
  scale_fill_manual(values = normal12,
                    name="Highest expressing\ncell type") +
  scale_colour_manual(values = normal12,
                    name="Highest expressing\ncell type") +
  guides(colour = guide_legend(override.aes = list(size=4))) +
  ylab("Difference in frequency (MAPK - Hypoxia)") +
  theme_classic() +
  theme(axis.text.x = element_blank(),
        axis.title.x = element_blank(),
        axis.ticks.x = element_blank(),
        axis.line.x = element_blank(),
        axis.text.y = element_text(size=14, colour="black"),
        axis.title.y = element_text(size=14),
        legend.text = element_text(size=14, color="black"),
        legend.title = element_text(size=14))

ggsave(diff_plot, filename="mapk_hypoxia_ligands.pdf",
       width=6, height=6, device=cairo_pdf())

diff_plot
```

```{r}

```

