---
title: "EMP Programs - MIX-seq analysis"
output: html_notebook
---

# Goal
Use the MIX-seq data set to find drugs that inhibit EMP programs in the cell lines assessed

# Dependencies
```{r}
library(Seurat)
library(tidyverse)
library(pheatmap)
library(ggridges)
```

# Load the data
I had previous downloaded the data and loaded it all into a single seurat object

```{r}
seurat <- readRDS("~/Data/single_cell_datasets/mix_seq/output/seurat_processed.rds")
```

Get an idea for how the data is organized
```{r}
head(seurat@meta.data)
```
And all the individual cell lines:
```{r}
table(seurat$singlet_ID)
```

Would be nice to also have this organized in two columns: 1) Line name, 2) cancer type
```{r}
seurat$CellLine_Full <- seurat$singlet_ID
```

```{r}
seurat@meta.data <- separate(seurat@meta.data, singlet_ID, into = c("CellLine", "Cancer"),
                             extra="merge")
```

```{r}
table(seurat$Cancer)
```

And then a unique sample ID (cell line + treatment) will be helpful down the road

```{r}
seurat$SampleID <- paste0(seurat$CellLine, "_", seurat$Drug)
```


## Get cell count summaries
The counts above include controls and all treatments. Let's get a good idea of counts for each condition
```{r}
dat <- seurat@meta.data %>%
  group_by(CellLine_Full, Drug) %>%
  summarise(Count = n())
```

```{r}
dat
```
Need to get an idea of cutoffs for including a cell line. Control treatments are the biggest deal here if I plan on defining a unique EMP program for each line. If that's the case, it may make sense to do 100-200 cell minimum
```{r}
dat %>% filter(Drug %in% c("DMSO", "Untreated") & Count >= 100) %>% nrow()
dat %>% filter(Drug %in% c("DMSO", "Untreated") & Count >= 200) %>% nrow()
```

# Naive gene set scoring
```{r}
emp_signature <- readLines("../output/conserved_emt_signature.txt")
```

```{r}
seurat <- AddModuleScore(seurat, features = list(emp_signature), name = "EMP_Sig")
```

```{r}
colnames(seurat@meta.data)[32] <- "EMP_Sig"
```


## Plot each cell line
Definitely excessive, but for now, I'll just do a quick function to make ridge plots for each cell line + each drug.

```{r}
seurat$Drug <- factor(seurat$Drug, 
                      levels = c("Untreated", "DMSO",
                                 unique(seurat$Drug)[-which(unique(seurat$Drug) %in% c("Untreated", "DMSO"))]))
```

```{r}
ridgeFun <- function(cell_line){
  dat <- filter(dat, CellLine == cell_line)
  
  dat_plot <- ggplot(dat, aes(x=EMP_Sig, y = Drug, fill=stat(x))) +
    geom_density_ridges_gradient(alpha=0.75, scale = 1.5) +
    scale_fill_viridis_c(name="EMP Sig", option="A") +
    xlab("EMP Signature") +
    theme_ridges()
  
  ggsave(dat_plot, filename=paste0("../figs/mix_seq_figures/gene_set_scoring/",
                                   cell_line, "EMP_Sig.pdf"),
         width=5.5, height=3.75)
  ggsave(dat_plot, filename=paste0("../figs/mix_seq_figures/gene_set_scoring/",
                                   cell_line, "EMP_Sig.png"),
         width=5.5, height=3.75)
}
```

```{r}
dat <- seurat@meta.data %>%
  group_by(CellLine, Drug) %>%
  summarise(Count = n())
good_lines <- dat %>% 
  filter(Drug %in% c("DMSO", "Untreated") & Count >= 100) %>%
  pull(CellLine)
```

```{r}
dat <- seurat@meta.data

dat_summary <- dat %>%
  group_by(SampleID) %>%
  summarise(Count = n())
good_samples <- filter(dat_summary, Count >= 50) %>% pull(SampleID)
dat <- filter(dat, SampleID %in% good_samples)

lapply(good_lines, ridgeFun)
```

So the plots are interesting. Some promising things. One, for example, is that the EMP signatures for most control cell liens actually have a little tail skewing off to the right, which would be consistent with the marker representing some EMP in the lines. Also, trametinib (MEK inhibitor) seems to fairly consistently increase signature levels. This actually the opposite of what we would expect. 

We can look at a sample cell line to try to see what's going on there:

### KNS81 example
```{r}
seurat_subset <- subset(seurat, subset = CellLine == "KNS81")
```

```{r}
RidgePlot(seurat_subset, group.by="Drug", features="EMP_Sig")
```


```{r}
tmp <- FindMarkers(seurat_subset, 
                   features = emp_signature[which(emp_signature %in% rownames(seurat_subset))], 
                   only.pos=F, 
                   logfc.threshold = 0, 
                   group.by="Drug", 
                   ident.1="Trametinib")
tmp %>% filter(p_val_adj <= 0.05) %>% arrange(desc(avg_log2FC))
```


I was originally worried that the increase in the EMP signature may end up being related to something like upregulation of the cytokeratins in the signature. Doesn't seem to be the case. In fact, it seems that there's up and down regulation of some pretty consistent Mes genes (eg. IL32 goes up, PMEPA1 goes down).

In the end, I think defining a unique EMP signature for each cell line and then assessing the drugs may reveal a cleaner association.


# ACTIONet on each cell line

Only look at cell lines with at least 100 cells in the control condition. Maaaay be cleaner to go to 200
```{r}
dat <- seurat@meta.data %>%
  group_by(CellLine, Drug) %>%
  summarise(Count = n())
good_lines <- dat %>% 
  filter(Drug %in% c("DMSO", "Untreated") & Count >= 100) %>%
  pull(CellLine) %>%
  unique() #Multiple lines in there >1 time
```

```{r}
runACTION <- function(sample){
  print(paste0("Processing sample: ", sample))
  #Subset seurat and convert to SCE
  seurat_subset <- subset(seurat, subset = CellLine == sample & Drug %in% c("DMSO", "Untreated"))
  sce <- Seurat::as.SingleCellExperiment(seurat_subset, assay="RNA")
  
  #Run ACTIONet
  ace <- reduce.ace(sce, reduced_dim = 25) #As far as I can tell, this relates to dimensionality of the data. 50 is default. Works well for full tissue, but I think it leads to representing some noise on these pure populations
  action <- run.ACTIONet(ace=ace, k_max=6,
                         min.cells.per.arch = 5) #Tried some tweaks and this works pretty well for me
  return(action)
}
```

```{r}
action_list <- lapply(good_lines, runACTION)
names(action_list) <- good_lines
```


```{r}
saveRDS(action_list, file="../output/ACTIONet_list_mixseq.rds")
```

```{r}
action_list <- readRDS("../output/ACTIONet_list_mixseq.rds")
```


# Build master metadata

```{r}
meta_list <- list()
for(i in 1:length(action_list)){
  meta_list[[i]] <- data.frame(Dim1 = action_list[[i]]$ACTIONet2D[,1],
                               Dim2 = action_list[[i]]$ACTIONet2D[,2],
                               Archetype_Membership = factor(action_list[[i]]$assigned_archetype))
  colnames(action_list[[i]]$archetype_footprint) <- paste0("A", 1:ncol(action_list[[i]]$archetype_footprint))
  meta_list[[i]] <- cbind(meta_list[[i]], action_list[[i]]$archetype_footprint)
}
names(meta_list) <- good_lines
```

# EMT Gene Set Scoring
## Load gene sets
```{r}
hallmarks <- fgsea::gmtPathways("~/Data/GeneLists/hallmark.genesets.v6.1.symbols.gmt")
hallmark_emt <- hallmarks["HALLMARK_EPITHELIAL_MESENCHYMAL_TRANSITION"]

go_terms <- fgsea::gmtPathways("~/Data/GeneLists/GOTerms.BP.v6.1.symbols.gmt")
go_emt <- go_terms["GO_EPITHELIAL_TO_MESENCHYMAL_TRANSITION"]

#cancerSEA
cancer_sea <- readRDS("~/Data/GeneLists/cancerSEA/cancerSEA_gene_sets.rds")
names(cancer_sea) <- paste0("CancerSEA_", names(cancer_sea))
cancer_sea_emt <- cancer_sea["CancerSEA_EMT"]

#My conserved upregulated list
conserved_mes <- list(readr::read_lines("~/Projects/emt_dynamics/output/conserved_upregulated_genes.csv")[-1])
names(conserved_mes) <- "Conserved_EMT_Up"

#My conserved downregulated list
conserved_epi <- list(readr::read_lines("~/Projects/emt_dynamics/output/conserved_downregulated_genes.csv")[-1])
names(conserved_epi) <- "Conserved_EMT_Down"

#dbEMT
dbEMT <- read.csv("~/Data/GeneLists/Zhao_dbEMT.csv")
dbEMT <- list(dbEMT$Symbol)
names(dbEMT) <- "dbEMT"

#Puram HNSCC pEMT
puram <- read.csv("~/Data/GeneLists/puram_hnscc_pEMT_signatures.csv")
puram <- list(unique(puram$Gene)) #All unique genes from their list of 50-gene NMF programs
names(puram) <- "Puram_pEMT"

#Taube Core EMT Up
taube_up <- list(read_lines("~/Data/GeneLists/Taube_Core_EMT_Up.txt"))
names(taube_up) <- "Taube_Core_EMT_Up"

#Kinker et al. EMT signatures
kinker <- read.csv("~/Data/GeneLists/Kinker_Tirosh_RHPs.csv")
kinker_1 <- list(Kinker_1 = kinker$EMT.I)
kinker_2 <- list(Kinker_2 = kinker$EMT.II)
kinker_3 <- list(Kinker_3 = kinker$EMT.III)

#EMP Signature
emp_signature <- readLines("../output/conserved_emt_signature.txt")
```

```{r}
emt_sets <- c(cancer_sea_emt, hallmark_emt,
             go_emt,  conserved_mes, dbEMT, puram, taube_up,
             kinker_1, kinker_2, kinker_3, list(emp_signature))
```

```{r}
names(emt_sets)
```

```{r}
names(emt_sets) <- c("CancerSEA", "Hallmark", "GO", 
                     "Cook", "dbEMT", "Puram_pEMT", "Taube",
                     "Kinker_I", "Kinker_II", "Kinker_III", "EMP_Sig")
```

## Scoring
We'll just run Seurat's AddModuleScore on each individual sample
```{r}
for(i in 1:length(good_lines)){
  seurat_subset <- subset(seurat, subset = CellLine == good_lines[i])
  seurat_subset <- CellCycleScoring(seurat_subset, s.features = cc.genes.updated.2019$s.genes,
                                    g2m.features = cc.genes.updated.2019$g2m.genes)
  # Gene set score each EMT gene set
  for(n in 1:length(emt_sets)){
    seurat_subset <- AddModuleScore(seurat_subset, features = emt_sets[n],
                                    name = names(emt_sets)[n])
  }
  
  seurat_subset$EMP_Sig <- NULL
  colnames(seurat_subset@meta.data)[32:42] <- gsub("1", "", colnames(seurat_subset@meta.data)[32:42]) #remove the stupid 1
  
  #Merge data into meta_list
  # Don't really know why the cell barcode order is different, but they all match *shrug*
  meta_list[[i]] <- cbind(meta_list[[i]], 
                        seurat_subset@meta.data[rownames(meta_list[[i]]),c(names(emt_sets),  
                                                   "S.Score", "G2M.Score", "Phase", 
                                                   "percent.mito", "nCount_RNA")])
}
```

```{r}
saveRDS(meta_list, file = "../output/master_metadata_mixseq.rds")
```

# Correlate w/ EMT
```{r}
getCor <- function(df){
  total_arch <- unique(paste0("A", unique(df$Archetype_Membership)))
  total_arch <- total_arch[order(total_arch)]
  cor_mat <- cor(df[,total_arch], df[,names(emt_sets)])
}

cor_list <- lapply(meta_list, getCor)
```

Need to get a list that matches source to sample
```{r}
dat <- seurat@meta.data %>% filter(CellLine %in% good_lines & Drug %in% c("Untreated", "DMSO"))
annotation <- unique(dat[,c("CellLine", "Cancer")])
cancer_list <- unique(annotation$Cancer)
```

Go source by source, making heatmap lists for each patient within that source

Oops, just meant to have gene sets ordered the same in each list

```{r}
gene_set_order <- rev(c("EMP_Sig", "Kinker_III", "dbEMT", "Kinker_I", "Cook", "Hallmark", "Puram_pEMT",
                    "Taube", "CancerSEA", "Kinker_II", "GO"))
```


```{r}
all_heatmaps <- list()
for(i in 1:length(cancer_list)){
  samples <- annotation %>% filter(Cancer == cancer_list[i]) %>% pull(CellLine)
  
  heat_list <- NULL
  for(n in 1:length(samples)){
    heat <- ComplexHeatmap::pheatmap(t(cor_list[[samples[n]]])[gene_set_order,],
                                   breaks=seq(-0.5,0.5, length.out=100),
                                   color = colorRampPalette(rev(brewer.pal(n = 7, name = "RdBu")))(100),
                                  cluster_rows=F,
                                  cluster_cols=T,
                                  clustering_method="ward.D2",
                                  treeheight_col = 0,
                                  fontsize = 10,
                                  legend = F,
                                  border_col="black")
    heat_list <- heat_list + heat
  }
  all_heatmaps[[i]] <- heat_list
  draw(all_heatmaps[[i]])
}
```


```{r}
pdf(file = "../figs/mix_seq_figures/archetype_emt_correlation/archetype_emt_correlation_Lung_CellLine.pdf",
    width=20, height=2.15)
all_heatmaps[[1]]
dev.off()

pdf(file = "../figs/mix_seq_figures/archetype_emt_correlation/archetype_emt_correlation_Stomach_CellLine.pdf",
    width=6, height=2.15)
all_heatmaps[[2]]
dev.off()

pdf(file = "../figs/mix_seq_figures/archetype_emt_correlation/archetype_emt_correlation_UrinaryTract_CellLine.pdf",
    width=4, height=2.15)
all_heatmaps[[3]]
dev.off()

pdf(file = "../figs/mix_seq_figures/archetype_emt_correlation/archetype_emt_correlation_Esophagus_CellLine.pdf",
    width=5, height=2.15)
all_heatmaps[[4]]
dev.off()

pdf(file = "../figs/mix_seq_figures/archetype_emt_correlation/archetype_emt_correlation_CNS_CellLine.pdf",
    width=8, height=2.15)
all_heatmaps[[5]]
dev.off()

pdf(file = "../figs/mix_seq_figures/archetype_emt_correlation/archetype_emt_correlation_Kidney_CellLine.pdf",
    width=6, height=2.15)
all_heatmaps[[6]]
dev.off()

pdf(file = "../figs/mix_seq_figures/archetype_emt_correlation/archetype_emt_correlation_Skin_CellLine.pdf",
    width=9, height=2.15)
all_heatmaps[[7]]
dev.off()

pdf(file = "../figs/mix_seq_figures/archetype_emt_correlation/archetype_emt_correlation_UpperAerodigestive_CellLine.pdf",
    width=3.75, height=2.15)
all_heatmaps[[8]]
dev.off()

pdf(file = "../figs/mix_seq_figures/archetype_emt_correlation/archetype_emt_correlation_LargeIntestine_CellLine.pdf",
    width=6.25, height=2.15)
all_heatmaps[[9]]
dev.off()

pdf(file = "../figs/mix_seq_figures/archetype_emt_correlation/archetype_emt_correlation_Ovary_CellLine.pdf",
    width=6.25, height=2.15)
all_heatmaps[[10]]
dev.off()

pdf(file = "../figs/mix_seq_figures/archetype_emt_correlation/archetype_emt_correlation_Endometrium_CellLine.pdf",
    width=3.75, height=2.15)
all_heatmaps[[11]]
dev.off()

pdf(file = "../figs/mix_seq_figures/archetype_emt_correlation/archetype_emt_correlation_Prostate_CellLine.pdf",
    width=3.75, height=2.15)
all_heatmaps[[12]]
dev.off()

pdf(file = "../figs/mix_seq_figures/archetype_emt_correlation/archetype_emt_correlation_BiliaryTract_CellLine.pdf",
    width=2.1, height=2.15)
all_heatmaps[[13]]
dev.off()

pdf(file = "../figs/mix_seq_figures/archetype_emt_correlation/archetype_emt_correlation_Breast_CellLine.pdf",
    width=7.4, height=2.15)
all_heatmaps[[14]]
dev.off()

pdf(file = "../figs/mix_seq_figures/archetype_emt_correlation/archetype_emt_correlation_Pleura_CellLine.pdf",
    width=3.75, height=2.15)
all_heatmaps[[15]]
dev.off()

pdf(file = "../figs/mix_seq_figures/archetype_emt_correlation/archetype_emt_correlation_Thyroid_CellLine.pdf",
    width=2.8, height=2.15)
all_heatmaps[[16]]
dev.off()

pdf(file = "../figs/mix_seq_figures/archetype_emt_correlation/archetype_emt_correlation_SoftTissue_CellLine.pdf",
    width=3.75, height=2.15)
all_heatmaps[[17]]
dev.off()

pdf(file = "../figs/mix_seq_figures/archetype_emt_correlation/archetype_emt_correlation_Stomach_CellLine.pdf",
    width=5.25, height=2.15)
all_heatmaps[[18]]
dev.off()

pdf(file = "../figs/mix_seq_figures/archetype_emt_correlation/archetype_emt_correlation_Liver_CellLine.pdf",
    width=2.1, height=2.15)
all_heatmaps[[19]]
dev.off()
```

# Get EMT programs
```{r}
cor_list_merge <- cor_list
for(i in 1:length(cor_list_merge)){
  rownames(cor_list_merge[[i]]) <- paste0(names(cor_list_merge)[i], "_", rownames(cor_list_merge[[i]]))
}
cor_list_merge <- do.call("rbind", cor_list_merge)
```

```{r}
cols <- paletteer::paletteer_c("scico::berlin", n=100)
```

```{r}
emt_cor_heatmap <- pheatmap::pheatmap(cor_list_merge[,gene_set_order],
                                      color = cols,
                                      breaks=seq(-0.75, 0.75, length.out=101),
                                      cluster_rows=T,
                                      cluster_cols=T,
                                      show_rownames=F,
                                      cutree_row=3,
                                      fontsize = 12,
                                      angle_col=45,
                                      clustering_method="ward.D2",
                                      #filename="../figs/mix_seq_figures/archetype_emt_correlation_all.png",
                                      width=4, height=8)

plot(emt_cor_heatmap$gtable)
```

```{r}
clusters <- as.data.frame(cutree(emt_cor_heatmap$tree_row, k=3))
colnames(clusters) <- "Cluster"
clusters$Program <- rownames(clusters)
table(clusters$Cluster)
```

Order from top to bottom
```{r}
unique(clusters$Cluster[emt_cor_heatmap$tree_row$order])
```

```{r}
emt_arch_high <- clusters %>% filter(Cluster==3) %>% pull(Program)
writeLines(emt_arch_high, file("../output/mixseq_emt_programs.txt"))
```

# Differential expression
```{r}
mtGenes <- grep('^MT-',rownames(seurat),value=TRUE)
```

# Functions
```{r}
##### Generic linear model function on pearson residuals of SCTransform model. Can change to test any other test
testGene <- function(gene, arch_score, seu){
  model_res <- summary(lm(seu[["SCT"]]@scale.data[gene,] ~ arch_score))
  res_table <- data.frame(pval = model_res$coefficients[2,4],
                          coef = model_res$coefficients[2,1],
                          Gene = gene)
  return(res_table)
}

##### Big function to get gene scores for each program of a given sample
testPrograms <- function(sampleID){
  print(paste0("Testing: ", sampleID))
  
  #Subset seurat
  print("Running SCTransform")
  seurat_subset <- subset(seurat, subset = CellLine == sampleID &
                            Drug %in% c("Untreated", "DMSO"))
  seurat_subset <- SCTransform(seurat_subset, vars.to.regress = "percent.mito")
  #Get gene residual variance data
  var_dat <- seurat_subset[["SCT"]]@meta.features
  var_dat$Gene <- rownames(var_dat)
  #Select top 2k genes with certain criteria
  top_genes <- var_dat %>%
    arrange(desc(sct.residual_variance)) %>%
    filter(sct.detection_rate >= 0.05) %>% # Require 5% detection rate
    pull(Gene)
  top_genes <- top_genes[1:2000] #Top 2k variable genes with those criteria
  top_genes <- top_genes[which(top_genes %in% rownames(seurat_subset[["SCT"]]@scale.data))]
  
  # Get archetype data
  meta <- meta_list[[sampleID]]
  arch_list <- unique(paste0("A", meta$Archetype_Membership))
  
  # Test genes for each archetype. 
  print("Testing for genes associated with archetype program activity")
  result_table <- list()
  for(i in 1:length(arch_list)){
    arch_score <- as.numeric(meta[,arch_list[i]])
    #Normalize archetype scores to have max = 1
    arch_score <- arch_score / max(arch_score)
    arch_res <- lapply(top_genes, testGene, arch_score = arch_score, seu = seurat_subset)
    arch_res <- do.call("rbind", arch_res)
    arch_res$padj <- p.adjust(arch_res$pval, method="BH")
    arch_res$Program <- paste0(sampleID, "_", arch_list[i])
    result_table[[i]] <- arch_res
  }
  result_table <- do.call("rbind", result_table)
}
```

```{r}
program_scores <- lapply(good_lines, testPrograms)

program_scores <- do.call("rbind", program_scores)
write.csv(program_scores, file="../output/archetype_program_scores_mixseq.csv", row.names = F)
```

```{r}
program_scores <- read.csv("../output/archetype_program_scores_mixseq.csv")
```

# Organize program scores
## -log10(padj) mat
For this matrix, I think I'll only keep significant changes w/ positive coefficients so that we don't get genes with hits in multiple programs without knowing that they're going in opposite directions.

```{r}
program_pval <- program_scores
program_pval <- filter(program_pval, coef > 0 & padj <= 0.05) # torn about doing only sig--should help
program_pval$coef <- NULL
program_pval$pval <- NULL
program_pval$padj <- -log10(program_pval$padj)
program_pval <- pivot_wider(program_pval, names_from=Program, values_from=padj)
program_pval <- as.data.frame(program_pval)
rownames(program_pval) <- program_pval$Gene
program_pval$Gene <- NULL
program_pval <- as.matrix(program_pval)
program_pval[is.na(program_pval)] <- 0
```

## Coefficient matrix
For this, I'll keep positive and negative values for the purpose of correlating programs

```{r}
program_coef <- program_scores
program_coef$padj <- NULL
program_coef$pval <- NULL
program_coef <- pivot_wider(program_coef, names_from=Program, values_from=coef)
program_coef <- as.data.frame(program_coef)
rownames(program_coef) <- program_coef$Gene
program_coef$Gene <- NULL
program_coef <- as.matrix(program_coef)
program_coef[is.na(program_coef)] <- 0
```

# Coefs of genes associated with EMT programs

Goal here: defining a signature of the strongest EMT-associated genes. For this, we only really care to consider the genes with pretty strong coefficients in the model. 

```{r}
#Get relevant genes
genes_keep <- program_scores %>%
  filter(Program %in% emt_arch_high &
           padj <= 0.05 &
           abs(coef) > 1) %>%
  pull(Gene) %>%
  unique()

emt_scores <- program_coef[genes_keep, emt_arch_high]

#emt_scores[emt_scores > 3] <- 3
#emt_scores[emt_scores < -3] <- -3
```

```{r}
emt_pval <- pheatmap::pheatmap(emt_scores,
                     color = cols,
                     breaks=seq(-3, 3, length.out=101),
                     clustering_method="ward.D2",
                     show_colnames=T,
                     angle_col = 45,
                     fontsize = 5,
                     show_rownames=F,
                     cutree_rows=3,
                     #cutree_cols = 4,
                     filename="../figs/mix_seq_figures/emt_program_coefficients_mixseq.png",
                     width=10, height=10)
```

# Test drugs for inhibition of EMP
## Define program signatures
```{r}
getGenes <- function(program){
  hits <- program_scores %>% filter(Program == program &
                                padj <= 0.05 &
                                coef >= 0.25) %>%
    pull(Gene)
}

program_genes <- lapply(unique(program_scores$Program), getGenes)
names(program_genes) <- unique(program_scores$Program)
```

```{r}
emp_programs <- program_genes[emt_arch_high]
```


A little wrangling to loop through each EMP program and score those genes in the seurat object


```{r}
annotation_df <- data.frame(ProgramFull = names(emp_programs))
annotation_df <- separate(annotation_df, ProgramFull,
                         into = c("CellLine", "Program"),
                         remove=F)
annotation_df$GoodName <- make.names(annotation_df$ProgramFull)#deal with stupid numeric cell lines
```

```{r}
emp_scores <- list() # will get meta data for each cell line
for(i in 1:nrow(annotation_df)){
  seurat_subset <- subset(seurat, subset = CellLine == annotation_df[i,"CellLine"])
  emp_genes <- emp_programs[[annotation_df[i,"ProgramFull"]]]
  seurat_subset <- AddModuleScore(seurat_subset, features = list(emp_genes),
                                  name = annotation_df[i,"GoodName"])
  #annoying
  colnames(seurat_subset@meta.data)[33] <- annotation_df[i,"GoodName"] #remove the stupid 1
  emp_scores[[i]] <- seurat_subset@meta.data
}
names(emp_scores) <- annotation_df$ProgramFull
```

## Plotting scripts
```{r}
for(i in 1:nrow(annotation_df)){
  program_name <- annotation_df[i, "GoodName"]
  dat_plot <- ggplot(emp_scores[[i]], aes_string(x=program_name, y = "Drug")) +
    geom_density_ridges_gradient(alpha=0.75, scale = 1.5, aes(fill=stat(x)),
                                 quantile_lines=T, quantiles=2) +
    scale_fill_viridis_c(name="EMP Sig", option="A") +
    xlab("EMP Signature") +
    theme_ridges()
  
  ggsave(dat_plot, 
         filename=paste0("../figs/mix_seq_figures/archetype_scoring/",
                         annotation_df[i, "ProgramFull"],
                         "_EMP.pdf"),
         width=5.5, height=3.75)
  
  ggsave(dat_plot, 
         filename=paste0("../figs/mix_seq_figures/archetype_scoring/",
                         annotation_df[i, "ProgramFull"],
                         "_EMP.png"),
         width=5.5, height=3.75)
}
```

## Single heatmap
```{r}
df_summary <- list()
for(i in 1:nrow(annotation_df)){
  dat <- emp_scores[[i]]
  program <- annotation_df$GoodName[i]
  
  # For the sake of this, let's just average out Untreated and DMSO
  dat$Drug <- as.character(dat$Drug)
  dat$Drug[dat$Drug %in% c("Untreated", "DMSO")] <- "Control"
  dat$Drug <- factor(dat$Drug,
                     levels = c("Control", 
                                unique(dat$Drug)[-which(unique(dat$Drug) == "Control")]))
  
  
  model_res <- summary(lm(dat[,program] ~ dat$Drug))
  model_res <- as.data.frame(coefficients(model_res))
  model_res <- model_res[2:nrow(model_res),c(1,4)]
  model_res$Drug <- substr(rownames(model_res), 9, nchar(rownames(model_res)))
  model_res$Program <- program
  
  df_summary[[i]] <- model_res
}

df_summary <- do.call("rbind", df_summary)
rownames(df_summary) <- 1:nrow(df_summary)
df_summary$padj <- p.adjust(df_summary$`Pr(>|t|)`, method="BH")
write.csv(df_summary, file="../output/mixseq_emp_drug_sensitivity.csv")
```

```{r}
#summary_mat <- df_summary %>% filter(padj <= 0.05) #Include all to distinguish between not tested and not significant
summary_mat <- df_summary[,c("Drug", "Program", "Estimate")]
```

```{r}
summary_mat <- pivot_wider(summary_mat, names_from = Program, values_from = Estimate)
```

```{r}
drug_names <- summary_mat$Drug
summary_mat <- as.matrix(summary_mat[,2:ncol(summary_mat)])
rownames(summary_mat) <- drug_names
```

```{r}
drug_heat <- pheatmap::pheatmap(summary_mat,
         color = colorRampPalette(RColorBrewer::brewer.pal(9, "PRGn"))(100),
         breaks = seq(-0.25, 0.25, length.out=101),
         na_col = "grey80",
         border_color = "black",
         clustering_method="ward.D2",
         angle_col = 45,
         fontsize_col = 7,
         filename="../figs/mix_seq_figures/drug_emp_coefs.png",
         width=13, height=4)
```

## Dotplot version
We'll steal the order from the heatmap though :P
```{r}
drug_levels <- rownames(summary_mat)[drug_heat$tree_row$order]
program_levels <- colnames(summary_mat)[drug_heat$tree_col$order]

df_summary$Drug <- factor(df_summary$Drug, levels=drug_levels)
df_summary$Program <- factor(df_summary$Program, levels = program_levels)

df_summary$log_qval <- -log10(df_summary$padj)
df_summary$log_qval[df_summary$log_qval > 10] <- 10 #hate these hacks
df_summary$Estimate[df_summary$Estimate > 0.2] <- 0.2
df_summary$Estimate[df_summary$Estimate < -0.2] <- -0.2
```


```{r}
drug_dot <- ggplot(df_summary, aes(x=Program, y=Drug)) +
  geom_point(aes(size = log_qval, fill = Estimate), color="black", shape=21) +
  scale_size("-log10(padj)", range = c(1,5)) +
  scale_fill_gradientn(colours = colorRampPalette(RColorBrewer::brewer.pal(9, "PRGn"))(100),
                       limits = c(-0.2, 0.2),
                       guide = guide_colorbar(ticks.colour = "black",
                                              frame.colour = "black"),
                       name = "Effect size") +
  ylab("") + xlab("") +
  theme_bw() +
  theme(axis.text.x = element_text(size=8, angle=45, hjust=1, color="black"),
        axis.text.y = element_text(size=12, color="black"))

ggsave(drug_dot, filename="../figs/mix_seq_figures/emt_program_coefficients_mixseq_dotplot.pdf",
       device = cairo_pdf(),
       width=14, height=5.2)
ggsave(drug_dot, filename="../figs/mix_seq_figures/emt_program_coefficients_mixseq_dotplot.png",
       width=14, height=5.2)
```

# PROGENy
Goal here is to relate the sensitivity of a cell line to a drug (eg. Trametinib) to its pathway activity. For example, in the tumours, half of the EMP programs had elevated EGFR/MEK signalling. About half of these cell lines have Trametinib-induced decreases in EMP activity. Goal is to see if these sensitive lines are the ones w/ high EGFR/MEK activity

```{r}
sample_list <- names(meta_list)
```

We'll only run PROGENy on the control conditions for now

```{r}
for(i in 1:length(sample_list)){
  cells_keep <- rownames(meta_list[[sample_list[i]]])
  seurat_subset <- subset(seurat, cells = cells_keep)
  seurat_subset <- SCTransform(seurat_subset, vars.to.regress="percent.mito")
  progeny_mat <- progeny(as.matrix(seurat_subset[["SCT"]]@scale.data),
                         top=500)
  meta_list[[i]] <- cbind(meta_list[[i]], progeny_mat)
}
```

```{r}
saveRDS(meta_list, file="../output/master_metadata_mixseq.rds")
```

## Activity ~ Program
Because each program has a different
```{r}
#pathways <- colnames(progeny_mat)
pathways <- c("Androgen", "EGFR", "Estrogen", "Hypoxia", "JAK-STAT", "MAPK", "NFkB", "p53", "PI3K", "TGFb", "TNFa", "Trail", "VEGF", "WNT")

testPathway <- function(pathway, program, df){
  score_norm <- df[,program] / max(df[,program])
  test_summary <- summary(lm(df[,pathway] ~ score_norm))
  test_res <- data.frame(Coef = test_summary$coefficients[2,1],
                         pval = test_summary$coefficients[2,4],
                         Pathway = pathway)
  return(test_res)
}

runPathwayTests <- function(sampleID){
  print(paste0("Testing: ", sampleID))
  df <- meta_list[[sampleID]]
  total_arch <- unique(paste0("A", unique(df$Archetype_Membership)))
  total_arch <- total_arch[order(total_arch)]
  
  pathway_test_res <- list()
  for(i in 1:length(total_arch)){
    test_summary <- lapply(pathways, testPathway, program = total_arch[i], df = df)
    test_summary <- do.call("rbind", test_summary)
    test_summary$padj <- p.adjust(test_summary$pval, method="BH")
    test_summary$Program <- paste0(sampleID, "_", total_arch[i])
    pathway_test_res[[i]] <- test_summary
  }
  
  pathway_test_res <- do.call("rbind", pathway_test_res)
}
```

```{r}
pathway_scores <- lapply(sample_list, runPathwayTests)
pathway_scores <- do.call("rbind", pathway_scores)
```

## PCA of coefficients
```{r}
pathway_coef <- pathway_scores
pathway_coef$padj <- NULL
pathway_coef$pval <- NULL
pathway_coef <- pivot_wider(pathway_coef, names_from=Pathway, values_from=Coef)
pathway_coef <- as.data.frame(pathway_coef)
rownames(pathway_coef) <- pathway_coef$Program
pathway_coef$Program <- NULL
pathway_coef <- as.matrix(pathway_coef)
rownames(pathway_coef) <- make.names(rownames(pathway_coef))
```

```{r}
pca_res <- prcomp(pathway_coef, scale=F)
```

```{r}
emt_program <- factor(rownames(pathway_coef) %in% make.names(emt_arch_high))
```


```{r}
pca1 <- fviz_pca_biplot(pca_res, geom.ind="point", axes=c(1,2),
                col.ind = emt_program, repel=T) +
  scale_color_manual(values=c('lightgrey', 'firebrick'))

pca2 <- fviz_pca_biplot(pca_res, geom.ind="point", axes=c(2,3),
                col.ind = emt_program, repel=T) +
  scale_color_manual(values=c('lightgrey', 'firebrick'))
pca1
pca2
```

```{r}
ggsave(pca1, filename="../figs/mix_seq_figures/progeny_EMP_pca.pdf",
       device = cairo_pdf(),
       width=5, height=3.5)
```


## Trametinib sensitivity

```{r}
tram_sensitivity <- df_summary %>%
  filter(Drug == "Trametinib")
tmp <- tram_sensitivity[match(rownames(pathway_coef), tram_sensitivity$Program), "Estimate"]
tmp[tmp < -0.1] <- -0.1
tmp[tmp > 0.1] <- 0.1
```

```{r}
pca1 <- fviz_pca_biplot(pca_res, geom.ind="point", axes=c(1,2),
                col.ind = tmp, repel=T) +
  scale_color_gradientn(colours=RColorBrewer::brewer.pal(9, "PRGn"))
pca1
```

```{r}
ggsave(pca1, filename="../figs/mix_seq_figures/progeny_Trametinib_coef.pdf",
       device = cairo_pdf(),
       width=5, height=3.5)
```



