---
title: "R Notebook"
output: html_notebook
---

# Goal
To get some idea of which stromal cell types may be driving EMP-associated expression programs 

# Dependencies
```{r}
library(Seurat)
library(nichenetr)
library(tidyverse)
library(ggrepel)
```

# Load EMP info
```{r}
program_scores <- read.csv("~/Projects/emt_programs/output/archetype_program_scores.csv")
```

```{r}
assessSig <- function(program){
  hits <- program_scores %>%
    filter(Program == program) %>%
    filter(padj <= 0.05 & coef > 1) %>%
    arrange(desc(padj)) %>% #Just so genes are actually ranked by significance in the list
    pull(Gene)#can add coefficient cutoffs if desired
  return(hits)
}

program_list <- unique(program_scores$Program)
program_genes <- lapply(program_list, assessSig)
names(program_genes) <- program_list
```

```{r}
emt_program_names <- readLines("~/Projects/emt_programs/output/emp_program_list_broad.txt")
emt_program <- program_genes[emt_program_names]
```

```{r}
meta_list <- readRDS("~/Projects/emt_programs/output/master_metadata.rds")
```

# Function through nichenet

A few nichenet components

```{r}
ligand_target_matrix = readRDS(url("https://zenodo.org/record/3260758/files/ligand_target_matrix.rds"))
lr_network = readRDS(url("https://zenodo.org/record/3260758/files/lr_network.rds"))
weighted_networks = readRDS(url("https://zenodo.org/record/3260758/files/weighted_networks.rds"))
weighted_networks_lr = weighted_networks$lr_sig %>% inner_join(lr_network %>% distinct(from,to), by = c("from","to"))
```


```{r}
programCheck <- function(sample_name){
  res <- grep(sample_name, emt_program_names, value=T)
  return(res)
}
```

```{r}
runNichenet <- function(sample){
  # Subset seurat by sample
  print(paste0("Testing sample: ", sample))
  seurat_subset <- subset(seurat, subset = Patient == sample)
  
  #Only want receiver cells to be archetypes - add EMP archetype info to seurat
  meta <- meta_list[[sample]]
  meta$ProgramFull <- paste0(sample, "_A", meta$Archetype_Membership)
  
  meta$EMP <- ifelse(meta$ProgramFull %in% emt_program_names, "_EMP", "")
  meta$CellType <- paste0("Cancer cells", meta$EMP)
  
  ####FIX CELL BARCODE TO REMOVE APPENDED UNIQUE NUMBER FROM COMBINING EPI DATA
  ####THIS IS GROSS
  char_remove <- paste0("_", gsub("^.*_", "", rownames(meta)))
  rownames(meta) <- gsub(paste0(unique(char_remove), "$"), "", rownames(meta))
  ####
  
  seurat_subset$CellType[match(rownames(meta), colnames(seurat_subset))] <- meta$CellType
  
  Idents(seurat_subset) <- seurat_subset$CellType
  
  ### Set up chance to exit if not enough cells
  ### In at least one case (Lung Wu P2) cases there isn't an EMP program
  if(sum(unique(meta$ProgramFull) %in% emt_program_names) == 0){
    print("No EMP program")
  } else {
  
  ### RUN NICHENET
  # Define receiver - EMP cells
  
  expressed_genes_receiver = get_expressed_genes("Cancer cells_EMP", seurat_subset, pct = 0.05,
                                               assay_oi = "RNA")
  background_expressed_genes = expressed_genes_receiver %>% .[. %in% rownames(ligand_target_matrix)]
  
  # Define senders as all possible cell types
  sender_celltypes = names(table(seurat_subset$CellType))[table(seurat_subset$CellType) > 20] #Don't consider sources w/ <20 cells
  list_expressed_genes_sender = sender_celltypes %>% unique() %>% 
    lapply(get_expressed_genes, seurat_subset, pct = 0.05, assay_oi = "RNA")
  expressed_genes_sender = list_expressed_genes_sender %>% unlist() %>% unique()
  
  ligands = lr_network %>% pull(from) %>% unique()
  receptors = lr_network %>% pull(to) %>% unique()

  expressed_ligands = intersect(ligands,expressed_genes_sender)
  expressed_receptors = intersect(receptors,expressed_genes_receiver)
  potential_ligands = lr_network %>% 
    filter(from %in% expressed_ligands & to %in% expressed_receptors) %>% 
    pull(from) %>% 
    unique()
  
  # Predict ligands
  ligand_activities = predict_ligand_activities(geneset = programs[[sample]], 
                                                background_expressed_genes = background_expressed_genes, 
                                                ligand_target_matrix = ligand_target_matrix, 
                                                potential_ligands = potential_ligands)
  
  ligand_activities$Sample <- sample
  
  ### ADD AVERAGE EXPRESSION
  ligands <- ligand_activities$test_ligand
  dat <- seurat_subset[["RNA"]]@data[ligands,]
  dat <- as.data.frame(t(as.matrix(dat)))
  dat$CellType <- seurat_subset$CellType
  
  dat <- dat %>% 
    pivot_longer(!CellType, names_to="Gene", values_to="Exp") %>%
    group_by(CellType, Gene) %>%
    summarise(pct = sum(Exp > 0) / length(Exp),
              AvgExp = mean(Exp)) %>%
    left_join(ligand_activities, by = c("Gene" = "test_ligand"))
  
  return(dat)
  }
}
```


# Colorectal - Lee et al - GSE144735 (SMC)

```{r}
seurat <- readRDS("../output/tumour_annotations/colorectal_lee_seurat.rds")
#Add some metadata to help post-merge
seurat$Cancer <-  "Colorectal"
seurat$Source <- "Lee"
seurat$Patient <- paste(seurat$Cancer, seurat$Source, seurat$Patient, sep="_")
```

```{r}
sample_list <- unique(seurat$Patient)
sample_program_list <- lapply(sample_list, programCheck)
names(sample_program_list) <- sample_list
sample_program_list <- sample_program_list[lengths(sample_program_list) > 0]
programs <- list()
for(i in 1:length(sample_program_list)){
  programs[[i]] <- emt_program[sample_program_list[[i]]]
  programs[[i]] <- unique(unlist(programs[[i]]))
}
names(programs) <- names(sample_program_list)
```

```{r}
colorectal_lee_smc <- lapply(names(programs), runNichenet)
names(colorectal_lee_smc) <- names(programs)
```

# Colorectal - Lee et al - GSE132465 (KUL)
```{r}
seurat <- readRDS("../output/tumour_annotations/colorectal_lee_KUL_seurat.rds")
#Add some metadata to help post-merge
seurat$Cancer <-  "Colorectal"
seurat$Source <- "Lee"
seurat$Patient <- paste(seurat$Cancer, seurat$Source, seurat$Patient, sep="_")
```


```{r}
sample_list <- unique(seurat$Patient)
sample_program_list <- lapply(sample_list, programCheck)
names(sample_program_list) <- sample_list
sample_program_list <- sample_program_list[lengths(sample_program_list) > 0]
programs <- list()
for(i in 1:length(sample_program_list)){
  programs[[i]] <- emt_program[sample_program_list[[i]]]
  programs[[i]] <- unique(unlist(programs[[i]]))
}
names(programs) <- names(sample_program_list)
```

```{r}
colorectal_lee_kul <- lapply(names(programs), runNichenet)
names(colorectal_lee_kul) <- names(programs)
```

# Colorectal - Uhlitz et al - Data direct from authors (preprint)
```{r}
seurat <- readRDS("../output/tumour_annotations/colorectal_Uhlitz_seurat.rds")
#Add some metadata to help post-merge
seurat$Cancer <-  "Colorectal"
seurat$Source <- "Uhlitz"
```

```{r}
seurat$Patient <- seurat$orig.ident
seurat$Patient <- paste(seurat$Cancer, seurat$Source, seurat$Patient, sep="_")
```

```{r}
sample_list <- unique(seurat$Patient)
sample_program_list <- lapply(sample_list, programCheck)
names(sample_program_list) <- sample_list
sample_program_list <- sample_program_list[lengths(sample_program_list) > 0]
programs <- list()
for(i in 1:length(sample_program_list)){
  programs[[i]] <- emt_program[sample_program_list[[i]]]
  programs[[i]] <- unique(unlist(programs[[i]]))
}
names(programs) <- names(sample_program_list)
```

```{r}
colorectal_uhlitz <- lapply(names(programs), runNichenet)
names(colorectal_uhlitz) <- names(programs)
```

# Colorectal - Qian et al - http://blueprint.lambrechtslab.org

```{r}
seurat <- readRDS("../output/tumour_annotations/colorectal_Qian_seurat.rds")
#Add some metadata to help post-merge
seurat$Cancer <-  "Colorectal"
seurat$Source <- "Qian"
```

```{r}
seurat$Patient <- seurat$orig.ident
seurat$Patient <- paste(seurat$Cancer, seurat$Source, seurat$Patient, sep="_")
```

```{r}
sample_list <- unique(seurat$Patient)
sample_program_list <- lapply(sample_list, programCheck)
names(sample_program_list) <- sample_list
sample_program_list <- sample_program_list[lengths(sample_program_list) > 0]
programs <- list()
for(i in 1:length(sample_program_list)){
  programs[[i]] <- emt_program[sample_program_list[[i]]]
  programs[[i]] <- unique(unlist(programs[[i]]))
}
names(programs) <- names(sample_program_list)
```

```{r}
colorectal_qian <- lapply(names(programs), runNichenet)
names(colorectal_qian) <- names(programs)
```

# Gastric - Sathe et al - https://dna-discovery.stanford.edu

```{r}
seurat <- readRDS("../output/tumour_annotations/gastric_sathe_seurat.rds")
#Add some metadata to help post-merge
seurat$Cancer <-  "Gastric"
seurat$Source <- "Sathe"
```

```{r}
seurat$Patient <- seurat$orig.ident
seurat$Patient <- paste(seurat$Cancer, seurat$Source, seurat$Patient, sep="_")
```

```{r}
sample_list <- unique(seurat$Patient)
sample_program_list <- lapply(sample_list, programCheck)
names(sample_program_list) <- sample_list
sample_program_list <- sample_program_list[lengths(sample_program_list) > 0]
programs <- list()
for(i in 1:length(sample_program_list)){
  programs[[i]] <- emt_program[sample_program_list[[i]]]
  programs[[i]] <- unique(unlist(programs[[i]]))
}
names(programs) <- names(sample_program_list)
```

```{r}
gastric_sathe <- lapply(names(programs), runNichenet)
names(gastric_sathe) <- names(programs)
```

# Lung - Kim et al - GSE131907

```{r}
seurat <- readRDS("../output/tumour_annotations/lung_kim_seurat.rds")
#Add some metadata to help post-merge
seurat$Cancer <-  "Lung"
seurat$Source <- "Kim"
seurat$Patient <- seurat$Sample
seurat$Patient <- paste(seurat$Cancer, seurat$Source, seurat$Patient, sep="_")
```


```{r}
sample_list <- unique(seurat$Patient)
sample_program_list <- lapply(sample_list, programCheck)
names(sample_program_list) <- sample_list
sample_program_list <- sample_program_list[lengths(sample_program_list) > 0]
programs <- list()
for(i in 1:length(sample_program_list)){
  programs[[i]] <- emt_program[sample_program_list[[i]]]
  programs[[i]] <- unique(unlist(programs[[i]]))
}
names(programs) <- names(sample_program_list)
```

```{r}
lung_kim <- lapply(names(programs), runNichenet)
names(lung_kim) <- names(programs)
```

# Lung - Lambrechts et al - E-MTAB-6149 & E-MTAB-6653

```{r}
seurat <- readRDS("../output/tumour_annotations/lung_lambrechts_seurat.rds")
#Add some metadata to help post-merge
seurat$Cancer <-  "Lung"
seurat$Source <- "Lambrechts"
```

```{r}
seurat$Patient <- seurat$orig.ident
seurat$Patient <- paste(seurat$Cancer, seurat$Source, seurat$Patient, sep="_")
```

```{r}
sample_list <- unique(seurat$Patient)
sample_program_list <- lapply(sample_list, programCheck)
names(sample_program_list) <- sample_list
sample_program_list <- sample_program_list[lengths(sample_program_list) > 0]
programs <- list()
for(i in 1:length(sample_program_list)){
  programs[[i]] <- emt_program[sample_program_list[[i]]]
  programs[[i]] <- unique(unlist(programs[[i]]))
}
names(programs) <- names(sample_program_list)
```

```{r}
lung_lambrechts <- lapply(names(programs), runNichenet)
names(lung_lambrechts) <- names(programs)
```

# Lung - Laughey et al - GSE123904

```{r}
seurat <- readRDS("../output/tumour_annotations/lung_laughney_seurat.rds")
#Add some metadata to help post-merge
seurat$Cancer <-  "Lung"
seurat$Source <- "Laughney"
seurat$Patient <- paste(seurat$Cancer, seurat$Source, seurat$Patient, sep="_")
```

```{r}
sample_list <- unique(seurat$Patient)
sample_program_list <- lapply(sample_list, programCheck)
names(sample_program_list) <- sample_list
sample_program_list <- sample_program_list[lengths(sample_program_list) > 0]
programs <- list()
for(i in 1:length(sample_program_list)){
  programs[[i]] <- emt_program[sample_program_list[[i]]]
  programs[[i]] <- unique(unlist(programs[[i]]))
}
names(programs) <- names(sample_program_list)
```

```{r}
lung_laughney <- lapply(names(programs), runNichenet)
names(lung_laughney) <- names(programs)
```

# Lung - Qian et al - http://blueprint.lambrechtslab.org

```{r}
seurat <- readRDS("../output/tumour_annotations/lung_qian_seurat.rds")
#Add some metadata to help post-merge
seurat$Cancer <-  "Lung"
seurat$Source <- "Qian"
seurat$Patient <- paste(seurat$Cancer, seurat$Source, seurat$Patient, sep="_")
```

```{r}
sample_list <- unique(seurat$Patient)
sample_program_list <- lapply(sample_list, programCheck)
names(sample_program_list) <- sample_list
sample_program_list <- sample_program_list[lengths(sample_program_list) > 0]
programs <- list()
for(i in 1:length(sample_program_list)){
  programs[[i]] <- emt_program[sample_program_list[[i]]]
  programs[[i]] <- unique(unlist(programs[[i]]))
}
names(programs) <- names(sample_program_list)
```

```{r}
lung_qian <- lapply(names(programs), runNichenet)
names(lung_qian) <- names(programs)
```

# Lung - Wu et al - GSE148071

```{r}
seurat <- readRDS("../output/tumour_annotations/lung_wu_seurat.rds")
#Add some metadata to help post-merge
seurat$Cancer <-  "Lung"
seurat$Source <- "Wu"
```

```{r}
seurat$Patient <- seurat$orig.ident
seurat$Patient <- paste(seurat$Cancer, seurat$Source, seurat$Patient, sep="_")
```

```{r}
sample_list <- unique(seurat$Patient)
sample_program_list <- lapply(sample_list, programCheck)
names(sample_program_list) <- sample_list
sample_program_list <- sample_program_list[lengths(sample_program_list) > 0]
programs <- list()
for(i in 1:length(sample_program_list)){
  programs[[i]] <- emt_program[sample_program_list[[i]]]
  programs[[i]] <- unique(unlist(programs[[i]]))
}
names(programs) <- names(sample_program_list)
```

```{r}
lung_wu <- lapply(names(programs), runNichenet)
names(lung_wu) <- names(programs)
```

# Ovarian - Geistlinger et al - GSE154600

```{r}
seurat <- readRDS("../output/tumour_annotations/ovarian_geistlinger_seurat.rds")
#Add some metadata to help post-merge
seurat$Cancer <-  "Ovarian"
seurat$Source <- "Geistlinger"
```

```{r}
seurat$Patient <- seurat$orig.ident
seurat$Patient <- paste(seurat$Cancer, seurat$Source, seurat$Patient, sep="_")
```

```{r}
sample_list <- unique(seurat$Patient)
sample_program_list <- lapply(sample_list, programCheck)
names(sample_program_list) <- sample_list
sample_program_list <- sample_program_list[lengths(sample_program_list) > 0]
programs <- list()
for(i in 1:length(sample_program_list)){
  programs[[i]] <- emt_program[sample_program_list[[i]]]
  programs[[i]] <- unique(unlist(programs[[i]]))
}
names(programs) <- names(sample_program_list)
```

```{r}
ovarian_geistlinger <- lapply(names(programs), runNichenet)
names(ovarian_geistlinger) <- names(programs)
```

# Ovarian - Qian et al - http://blueprint.lambrechtslab.org

```{r}
seurat <- readRDS("../output/tumour_annotations/ovarian_qian_seurat.rds")
#Add some metadata to help post-merge
seurat$Cancer <-  "Ovarian"
seurat$Source <- "Qian"
seurat$Patient <- paste(seurat$Cancer, seurat$Source, seurat$Patient, sep="_")
```

```{r}
sample_list <- unique(seurat$Patient)
sample_program_list <- lapply(sample_list, programCheck)
names(sample_program_list) <- sample_list
sample_program_list <- sample_program_list[lengths(sample_program_list) > 0]
programs <- list()
for(i in 1:length(sample_program_list)){
  programs[[i]] <- emt_program[sample_program_list[[i]]]
  programs[[i]] <- unique(unlist(programs[[i]]))
}
names(programs) <- names(sample_program_list)
```

```{r}
ovarian_qian <- lapply(names(programs), runNichenet)
names(ovarian_qian) <- names(programs)
```


# PDAC - Steele et al - GSE155698

```{r}
seurat <- readRDS("../output/tumour_annotations/pdac_steele_seurat.rds")
#Add some metadata to help post-merge
seurat$Cancer <-  "PDAC"
seurat$Source <- "Steele"
seurat$Patient <- gsub("PDAC_", "", seurat$Sample)

seurat$Patient <- paste(seurat$Cancer, seurat$Source, seurat$Patient, sep="_")
```


```{r}
sample_list <- unique(seurat$Patient)
sample_program_list <- lapply(sample_list, programCheck)
names(sample_program_list) <- sample_list
sample_program_list <- sample_program_list[lengths(sample_program_list) > 0]
programs <- list()
for(i in 1:length(sample_program_list)){
  programs[[i]] <- emt_program[sample_program_list[[i]]]
  programs[[i]] <- unique(unlist(programs[[i]]))
}
names(programs) <- names(sample_program_list)
```

```{r}
pdac_steele <- lapply(names(programs), runNichenet)
names(pdac_steele) <- names(programs)
```

# Breast - Wu et al - ENA PRJEB35405

```{r}
seurat <- readRDS("../output/tumour_annotations/breast_wu_seurat.rds")
#Add some metadata to help post-merge
seurat$Cancer <-  "Breast"
seurat$Source <- "Wu"
seurat$Patient <- paste(seurat$Cancer, seurat$Source, seurat$Patient, sep="_")
```

```{r}
sample_list <- unique(seurat$Patient)
sample_program_list <- lapply(sample_list, programCheck)
names(sample_program_list) <- sample_list
sample_program_list <- sample_program_list[lengths(sample_program_list) > 0]
programs <- list()
for(i in 1:length(sample_program_list)){
  programs[[i]] <- emt_program[sample_program_list[[i]]]
  programs[[i]] <- unique(unlist(programs[[i]]))
}
names(programs) <- names(sample_program_list)
```

```{r}
breast_wu <- lapply(names(programs), runNichenet)
names(breast_wu) <- names(programs)
```

# Breast - Qian et al - http://blueprint.lambrechtslab.org

```{r}
seurat <- readRDS("../output/tumour_annotations/breast_qian_seurat.rds")
#Add some metadata to help post-merge
seurat$Cancer <-  "Breast"
seurat$Source <- "Qian"
seurat$Patient <- paste(seurat$Cancer, seurat$Source, seurat$Patient, sep="_")
```

```{r}
sample_list <- unique(seurat$Patient)
sample_program_list <- lapply(sample_list, programCheck)
names(sample_program_list) <- sample_list
sample_program_list <- sample_program_list[lengths(sample_program_list) > 0]
programs <- list()
for(i in 1:length(sample_program_list)){
  programs[[i]] <- emt_program[sample_program_list[[i]]]
  programs[[i]] <- unique(unlist(programs[[i]]))
}
names(programs) <- names(sample_program_list)
```

```{r}
breast_qian <- lapply(names(programs), runNichenet)
names(breast_qian) <- names(programs)
```

# Breast - Bassez et al - http://biokey.lambrechtslab.org/

## Cohort 1

```{r}
seurat <- readRDS("../output/tumour_annotations/breast_bassez_cohort1_seurat.rds")
#Add some metadata to help post-merge
seurat$Cancer <-  "Breast"
seurat$Source <- "Bassez"
seurat$Patient <- paste(seurat$Cancer, seurat$Source, seurat$Patient, sep="_")
```

```{r}
sample_list <- unique(seurat$Patient)
sample_program_list <- lapply(sample_list, programCheck)
names(sample_program_list) <- sample_list
sample_program_list <- sample_program_list[lengths(sample_program_list) > 0]
programs <- list()
for(i in 1:length(sample_program_list)){
  programs[[i]] <- emt_program[sample_program_list[[i]]]
  programs[[i]] <- unique(unlist(programs[[i]]))
}
names(programs) <- names(sample_program_list)
```

```{r}
breast_bassez_cohort1 <- lapply(names(programs), runNichenet)
names(breast_bassez_cohort1) <- names(programs)
```

## Cohort 2

```{r}
seurat <- readRDS("../output/tumour_annotations/breast_bassez_cohort2_seurat.rds")
#Add some metadata to help post-merge
seurat$Cancer <-  "Breast"
seurat$Source <- "Bassez"
seurat$Patient <- paste(seurat$Cancer, seurat$Source, seurat$Patient, sep="_")
```

```{r}
sample_list <- unique(seurat$Patient)
sample_program_list <- lapply(sample_list, programCheck)
names(sample_program_list) <- sample_list
sample_program_list <- sample_program_list[lengths(sample_program_list) > 0]
programs <- list()
for(i in 1:length(sample_program_list)){
  programs[[i]] <- emt_program[sample_program_list[[i]]]
  programs[[i]] <- unique(unlist(programs[[i]]))
}
names(programs) <- names(sample_program_list)
```

```{r}
breast_bassez_cohort2 <- lapply(names(programs), runNichenet)
names(breast_bassez_cohort2) <- names(programs)
```

# SCC - Ji et al - GSE144236

```{r}
seurat <- readRDS("../output/tumour_annotations/scc_ji_seurat.rds")
#Add some metadata to help post-merge
seurat$Cancer <-  "SCC"
seurat$Source <- "Ji"
seurat$Patient <- paste(seurat$Cancer, seurat$Source, seurat$Patient, sep="_")
```

```{r}
sample_list <- unique(seurat$Patient)
sample_program_list <- lapply(sample_list, programCheck)
names(sample_program_list) <- sample_list
sample_program_list <- sample_program_list[lengths(sample_program_list) > 0]
programs <- list()
for(i in 1:length(sample_program_list)){
  programs[[i]] <- emt_program[sample_program_list[[i]]]
  programs[[i]] <- unique(unlist(programs[[i]]))
}
names(programs) <- names(sample_program_list)
```

```{r}
scc_ji <- lapply(names(programs), runNichenet)
names(scc_ji) <- names(programs)
```

# Nasopharyngeal Carcinoma - Chen et al - GSE150430

```{r}
seurat <- readRDS("../output/tumour_annotations/npc_chen_seurat.rds")
#Add some metadata to help post-merge
seurat$Cancer <-  "NPC"
seurat$Source <- "Chen"
seurat$Patient <- paste(seurat$Cancer, seurat$Source, seurat$Patient, sep="_")
```

```{r}
sample_list <- unique(seurat$Patient)
sample_program_list <- lapply(sample_list, programCheck)
names(sample_program_list) <- sample_list
sample_program_list <- sample_program_list[lengths(sample_program_list) > 0]
programs <- list()
for(i in 1:length(sample_program_list)){
  programs[[i]] <- emt_program[sample_program_list[[i]]]
  programs[[i]] <- unique(unlist(programs[[i]]))
}
names(programs) <- names(sample_program_list)
```

```{r}
npc_chen <- lapply(names(programs), runNichenet)
names(npc_chen) <- names(programs)
```

# Combine data
```{r}
niche_res <- c(colorectal_lee_smc, colorectal_lee_kul, colorectal_uhlitz, colorectal_qian,
                 lung_kim, lung_laughney, lung_lambrechts, lung_qian, pdac_steele,
                 scc_ji, breast_wu, breast_qian, ovarian_geistlinger, ovarian_qian,
               gastric_sathe, breast_bassez_cohort1, breast_bassez_cohort2, npc_chen,
               lung_wu)

#Lung_Wu_P2 is null because it had no cells as cancer_EMP

niche_res <- niche_res[-which(names(niche_res)=="Lung_Wu_P2")]

niche_res <- lapply(niche_res, as.data.frame)
niche_res <- do.call("bind_rows", niche_res)

#when we combined epithelial cells, we omitted cells annotated as "Epithelial" that were part of a cluster that was very clearly not epithelial (determined by the "epithelial" annotation representing a very small fraction of the cluster)

#These stray epithelial cells were included as senders in the analysis. Because these strays spanned multiple clusters and were rare, it makes the most sense to remove them. Fortunately they are the only cells left annotated as "Epithelial_cells", so they can be filtered easily
niche_res <- filter(niche_res, CellType != "Epithelial_cells")
```

```{r}
saveRDS(niche_res, file="../output/nichenet_res.rds")
```

```{r}
niche_res <- readRDS("../output/nichenet_res.rds")
```

# Explore a bit
```{r}
pearson_dist_plot <- niche_res %>%
  select(c(Sample, Gene, pearson)) %>%
  unique() %>%
  ggplot(aes(Sample, pearson)) +
  geom_boxplot(outlier.size = 0, fill="firebrick", alpha=0.5) +
  xlab("") + ylab("Pearson") +
  theme_classic() +
  theme(axis.text.x = element_text(angle=45, hjust=1, colour="black", size=8),
        axis.text.y = element_text(size=10, colour="black"))
ggsave(pearson_dist_plot, filename="../figs/nichenet_pearson_dist.pdf",
       width=15, height=4.75, device=cairo_pdf())
```

```{r}
pearson_dist_plot
```
Given the variation in the distribution, I think something like top 10-20 per sample seems reasonable. Let's go with 10

```{r}
good_ligands <- niche_res %>%
  select(c("Sample", "pearson", "Gene")) %>%
  unique() %>%
  group_by(Sample) %>%
  top_n(10, pearson) %>%
  ungroup() %>%
  group_by(Gene) %>%
  summarise(Count = length(Sample)) %>%
  arrange(desc(Count)) %>%
  mutate(Rank = 1:nrow(.))
good_ligands
```

```{r}
normal12 <- c("#ebac23", #EMP
              "#008cf9",#Cancer
              "#b80058", #Fib
              "#B24502", #Smooth muscle
              "#00bbad", #Endo
              "#d163e6", #Macro
              "#FF9287", #Mono
              "#006e00", #B
              "#5954D6", #T
              "#00C6F8", #NK
              "#878500", #DC
              "#00A76C", #Platelets
              "#bdbdbd") #Unkown
```


```{r}
good_ligands$Label <- ""
good_ligands$Label[1:25] <- good_ligands$Gene[1:25]


## Add color by highest expressing cell type
niche_summary <- niche_res %>%
  filter(Gene %in% unique(good_ligands$Gene)) %>%
  group_by(CellType, Gene) %>%
  summarise(Exp = mean(AvgExp, na.rm = T),
            Pct = mean(pct, na.rm=T)) %>%
  ungroup() %>%
  group_by(Gene) %>%
  filter(Exp == max(Exp))

merge_dat <- left_join(good_ligands, niche_summary, by="Gene")
merge_dat$CellType <- factor(merge_dat$CellType, 
                             levels = c("Cancer cells_EMP", "Cancer cells", "Fibroblasts",
                                        "Smooth_muscle_cells", "Endothelial_cells", 
                                        "Macrophage", "Monocyte", "B_cell", 
                                        "T_cells", "NK_cell", "DC", "Platelets", "Unknown"))


ligand_freq_plot <- ggplot(merge_dat, aes(Rank, Count, label=Label)) +
  geom_point(size=1, color="black") +
  geom_text_repel(aes(color=CellType),
                  max.overlaps = 1000,
                  min.segment.length=0,
                  direction="y",
                  nudge_x=125,
                  size=3.25,
                  hjust=0,
                  segment.size=0.2,
                  segment.alpha=0.75,
                  segment.color="black") +
  xlab("Ranked Ligands") + ylab("Top ligand frequency (/266)") + 
  scale_y_continuous(breaks=c(0,20, 40, 60, 80, 100), limits=c(0,100)) +
  scale_color_manual(values=normal12) +
  theme_classic() +
  theme(axis.text.x = element_blank(),
        axis.ticks.x = element_blank(),
        axis.text.y = element_text(size=10, color="black"),
        axis.title.y = element_text(size=12))

ggsave(ligand_freq_plot, filename="../figs/nichenet_top_ligand_freq.pdf",
       width=5, height=3.75, device=cairo_pdf())
```

```{r}
ligand_freq_plot
```

# DotPlot Summary

Essentially recreating the standard Seurat dotplot 

```{r}
ligands <- good_ligands$Gene[1:50]
```


```{r}
niche_summary <- niche_res %>%
  filter(Gene %in% ligands) %>%
  group_by(CellType, Gene) %>%
  summarise(Exp = mean(AvgExp, na.rm = T),
            Pct = mean(pct, na.rm=T))
```


I'm going to make a dotplot, but I'd like the ligands/cell types to be ordered in a visually appealing way. I'll make a clustered heatmap, steal the order, and then use it on this
```{r}
niche_tmp <- niche_summary %>%
  select(c("CellType", "Gene", "Exp")) %>%
  pivot_wider(names_from="Gene", values_from="Exp")
niche_mat <- as.matrix(niche_tmp[,2:ncol(niche_tmp)])
rownames(niche_mat) <- niche_tmp$CellType

niche_heat <- pheatmap::pheatmap(niche_mat,
                                 silent=T,
                                 clustering_method = "ward.D2")
```


```{r}
#Aesthetic things
niche_dot <- niche_summary
niche_dot$Exp[niche_dot$Exp > 1] <- 1
niche_dot$Pct[niche_dot$Pct > 0.5] <- 0.5
niche_dot$Pct <- niche_dot$Pct * 100
niche_dot$CellType <- factor(niche_dot$CellType,
                             levels = rownames(niche_mat)[niche_heat$tree_row$order])
niche_dot$Gene <- factor(niche_dot$Gene,
                             levels = colnames(niche_mat)[niche_heat$tree_col$order])

#Plot
dot <- ggplot(niche_dot, aes(x=Gene, y=CellType)) +
  geom_point(aes(size = Pct, fill = Exp), color="black", shape=21) +
  scale_size("% Exp", range = c(0,6)) +
  scale_fill_gradientn(colours = viridisLite::mako(100),
                       limits = c(0,1),
                       guide = guide_colorbar(ticks.colour = "black",
                                              frame.colour = "black"),
                       name = "Avg Exp") +
  ylab("") + xlab("") +
  theme_bw() +
  theme(axis.text.x = element_text(size=10, angle=45, hjust=1, color="black"),
        axis.text.y = element_text(size=12, color="black"))

#Save
ggsave(dot, filename="../figs/nichnet_dotplot.pdf",
       device = cairo_pdf(),
       width=12, height=4.25)
ggsave(dot, filename="../figs/nichnet_dotplot.png",
       width=12, height=4.25) #width=12
```

# MAPK vs. Hypoxia EMP

## Run PCA first
## Prep data
```{r}
sample_list <- names(meta_list)
```

```{r}
#pathways <- colnames(progeny_mat)
pathways <- c("Androgen", "EGFR", "Estrogen", "Hypoxia", "JAK-STAT", "MAPK", "NFkB", "p53", "PI3K", "TGFb", "TNFa", "Trail", "VEGF", "WNT")

testPathway <- function(pathway, program, df){
  score_norm <- df[,program] / max(df[,program])
  test_summary <- summary(lm(df[,pathway] ~ score_norm))
  test_res <- data.frame(Coef = test_summary$coefficients[2,1],
                         pval = test_summary$coefficients[2,4],
                         Pathway = pathway)
  return(test_res)
}

runPathwayTests <- function(sampleID){
  print(paste0("Testing: ", sampleID))
  df <- meta_list[[sampleID]]
  total_arch <- unique(paste0("A", unique(df$Archetype_Membership)))
  total_arch <- total_arch[order(total_arch)]
  
  pathway_test_res <- list()
  for(i in 1:length(total_arch)){
    test_summary <- lapply(pathways, testPathway, program = total_arch[i], df = df)
    test_summary <- do.call("rbind", test_summary)
    test_summary$padj <- p.adjust(test_summary$pval, method="BH")
    test_summary$Program <- paste0(sampleID, "_", total_arch[i])
    pathway_test_res[[i]] <- test_summary
  }
  
  pathway_test_res <- do.call("rbind", pathway_test_res)
}
```

```{r}
pathway_scores <- lapply(sample_list, runPathwayTests)
pathway_scores <- do.call("rbind", pathway_scores)
```

```{r}
pathway_coef <- pathway_scores
pathway_coef$padj <- NULL
pathway_coef$pval <- NULL
pathway_coef <- pivot_wider(pathway_coef, names_from=Pathway, values_from=Coef)
pathway_coef <- as.data.frame(pathway_coef)
rownames(pathway_coef) <- pathway_coef$Program
pathway_coef$Program <- NULL
pathway_coef <- as.matrix(pathway_coef)
```

```{r}
pca_res <- prcomp(pathway_coef, scale=F)
```

```{r}
emp_programs <- readLines("../output/emp_program_list_broad.txt")
emt_program <- factor(rownames(pathway_coef) %in% emp_programs)
```

```{r}
pca1 <- fviz_pca_biplot(pca_res, geom.ind="point", axes=c(1,2),
                col.ind = emt_program, repel=T) +
  scale_color_manual(values=c('lightgrey', 'firebrick'))

pca1
```

## Compare MAPK vs. JAK-STAT
```{r}
mapk_programs <- rownames(pca_res$x)[rownames(pca_res$x) %in% emp_programs &
                                       pca_res$x[,2] < -1 &
                                       pca_res$x[,1] > 0]
mapk_programs <- substr(mapk_programs, 1, nchar(mapk_programs)-3) %>% unique()

stat_programs <- rownames(pca_res$x)[rownames(pca_res$x) %in% emp_programs &
                                       pca_res$x[,2] > 1 &
                                       pca_res$x[,1] > 0]
stat_programs <- substr(stat_programs, 1, nchar(stat_programs)-3) %>% unique()
```

```{r}
mapk_ligands <- niche_res %>%
  filter(Sample %in% mapk_programs) %>%
  select(c("Sample", "pearson", "Gene")) %>%
  unique() %>%
  group_by(Sample) %>%
  top_n(20, pearson) %>% #results are actually a bit cleaner if you go with top 10
  ungroup()



mapk_ligand_count <- mapk_ligands %>% group_by(Gene) %>%
  summarise(Count = length(Sample)) %>%
  arrange(desc(Count)) %>%
  mutate(Rank = 1:nrow(.))
mapk_ligand_count$Count <- mapk_ligand_count$Count / length(mapk_programs)
#The above line accounts for differences in the number of MAPK and STAT-associated programs. Convert to proportion


stat_ligands <- niche_res %>%
  filter(Sample %in% stat_programs) %>%
  select(c("Sample", "pearson", "Gene")) %>%
  unique() %>%
  group_by(Sample) %>%
  top_n(20, pearson) %>% #results are actually a bit cleaner if you go with top 10
  ungroup()



stat_ligand_count <- stat_ligands %>% group_by(Gene) %>%
  summarise(Count = length(Sample)) %>%
  arrange(desc(Count)) %>%
  mutate(Rank = 1:nrow(.))
stat_ligand_count$Count <- stat_ligand_count$Count / length(stat_programs) 
```

```{r}
#common_ligands <- intersect(mapk_ligand_count, stat_ligand_count)

merge_dat <- left_join(mapk_ligand_count, stat_ligand_count, by="Gene")
merge_dat <- na.omit(merge_dat)
merge_dat$CountDiff <- merge_dat$Count.x - merge_dat$Count.y #The difference in proportion
merge_dat <- arrange(merge_dat, desc(CountDiff))
merge_dat$Rank <- 1:nrow(merge_dat)

```

## Plot
Add colour by cell type with highest expression maybe??
```{r}
mapk_ligands <- c(merge_dat %>% filter(CountDiff > 0.075) %>% pull(Gene)) 
stat_ligands <- c(merge_dat %>% filter(CountDiff < -0.075) %>% pull(Gene))

merge_dat$mapk_Label <- ""
merge_dat$mapk_Label[match(mapk_ligands, merge_dat$Gene)] <- mapk_ligands

merge_dat$stat_Label <- ""
merge_dat$stat_Label[match(stat_ligands, merge_dat$Gene)] <- stat_ligands

# Get expression info for ligands
niche_summary <- niche_res %>%
  filter(Gene %in% unique(merge_dat$Gene)) %>%
  group_by(CellType, Gene) %>%
  summarise(Exp = mean(AvgExp, na.rm = T),
            Pct = mean(pct, na.rm=T)) %>%
  ungroup() %>%
  group_by(Gene) %>%
  filter(Exp == max(Exp))

merge_dat <- left_join(merge_dat, niche_summary, by="Gene")
merge_dat$CellType <- factor(merge_dat$CellType, 
                             levels = c("Cancer cells_EMP", "Cancer cells", "Fibroblasts",
                                        "Smooth_muscle_cells", "Endothelial_cells", 
                                        "Macrophage", "Monocyte", "B_cell", 
                                        "T_cells", "NK_cell", "DC", "Platelets", "Unknown"))

# Plot
diff_plot <- ggplot(merge_dat, aes(x=Rank, y=CountDiff)) +
  geom_point(aes(fill=CellType, size=Exp), colour="black", shape=21, alpha=0.6) +
  geom_hline(yintercept = c(-0.075, 0.075), linetype=2) +
  geom_text_repel(aes(label = mapk_Label, color=CellType),
                  nudge_x = 50,
                  direction = "y",
                  force=0.5,
                  ylim = c(0.075,NA),
                  hjust= 0,
                  size=3,
                  segment.size = 0.2,
                  segment.color="black") +
  geom_text_repel(aes(label = stat_Label, color=CellType),
                  nudge_x = -50,
                  force=0.5,
                  direction = "y",
                  ylim = c(NA,-0.075),
                  hjust= 1,
                  size=3,
                  segment.size = 0.2,
                  segment.color="black") +
  scale_y_continuous(expand = c(0, 0), breaks=c(-0.3, -0.2, -0.1, 0, 0.1, 0.2, 0.3), limits=c(-0.3, 0.3)) +
  scale_fill_manual(values = normal12,
                    name="Highest expressing\ncell type") +
  scale_colour_manual(values = normal12,
                    name="Highest expressing\ncell type") +
  guides(colour = guide_legend(override.aes = list(size=4))) +
  ylab("Difference in program frequency") +
  theme_classic() +
  theme(axis.text.x = element_blank(),
        axis.title.x = element_blank(),
        axis.ticks.x = element_blank(),
        axis.line.x = element_blank(),
        axis.text.y = element_text(size=10, colour="black"),
        axis.title.y = element_text(size=12),
        legend.text = element_text(size=10, color="black"),
        legend.title = element_text(size=12))

diff_plot
```





